---
title: "Adding a new mzMl file to the database"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This Quarto document presents the processing workflow for a new mzML (raw) file
corresponding to a herbacetin experiment. It describes the steps required to add
the experiment to the SQL database.


##Load required libraries

```{r}
#| message: false

#' Load required libraries
library(MsExperiment)
library(Spectra)
library(xcms)
library(readxl)
library(pander)
library(pheatmap)
library(S4Vectors)
library(devtools)
library(DBI)
library(RSQLite)
library(here)
library(dplyr)
library(tidyr)
library(IRanges)
library(MSnbase)
rm(list = c("conv.CSPP_SQL"))
#` Uploading RDynLib: Load all functions of the RDynlib directory
load_all()
register(SerialParam())
```

Import the mzml file  
```{r}
mzml_file <- here("data","BeTio_herbacetin1327_F2T3.mzML")
```

#Centroiding 
```{r}
sp <- Spectra(mzml_file)
a_c <- Spectra::pickPeaks(sp, halfWindowSize = 10L, k = 2L, threshold = 0.33,
                          msLevel. = 1L)

```

Next we export the centroided file
```{r}
export(
  a_c,
  backend = MsBackendMzR(),
  file = here("data","BeTio_herbacetin1327_F2T3_centroided.mzML"),
  format = "mzML",
  copy = FALSE
)
```
#Convert the mzml file to an MSExperiment object

Convert the mzML file to an MSExperiment object and import 
the corresponding sample data
```{r}
mzml_file_c <- here("data","BeTio_herbacetin1327_F2T3_centroided.mzML")
sample_data_ftms <- DataFrame(
    sample_id = "S1",
    sample_name = "Sample_1"
)
ftms <- readMsExperiment(
    mzml_file_c,
    sampleData = sample_data_ftms
)
ftms
```

colors assignment for the heatmap visualization
```{r}
sampleData(ftms)
col_sample_ftms <- "blue"
col_sample_ftms_id <- c(S1 = "blue")
col_sample_ftms <- "blue"
```

## Base peak chromatogram
```{r}
bpc0 <- chromatogram(ftms, aggregationFun = "max")

plot(bpc0, col = col_sample_ftms)
grid()
```

## Filter by mass to get the feature range of herbacetin
```{r}
ftms <- filterSpectra(ftms, filterMzRange, mz = c(625.0, 625.4), msLevel. = 1L)

table(msLevel(spectra(ftms)))
```

## Filter by rt to get the retention time range of herbacetin
```{r}
ftms <- filterSpectra(ftms, filterRt, c(690, 790))

table(msLevel(spectra(ftms)))
```

## Remove empty spectra 
```{r}
ftms <- filterSpectra(ftms, filterEmptySpectra)

table(msLevel(spectra(ftms)))
```

#Visualizing the base peak chromatogram again:
```{r}
bpc0 <- chromatogram(ftms, aggregationFun = "max")

plot(bpc0, col = col_sample_ftms)
grid()
```

#Peak detection
```{r}
ftms <- findChromPeaks(ftms, CentWaveParam(peakwidth = c(20, 80),
                                            integrate = 2, snthresh = 2))
```
```{r}
pks <- chromPeaks(ftms)
quantile(unname(pks[, "rtmax"] - pks[, "rtmin"]))
quantile(unname(pks[, "mzmax"] - pks[, "mzmin"]))
```


Change the backend to MsBackendMemory for further analysis
```{r}
ftms_sub <- ftms
## Change the backend to in-memory backend
ftms_sub@spectra <- setBackend(spectra(ftms_sub), MsBackendMemory())
## Apply any processing steps
ftms_sub@spectra <- applyProcessing(spectra(ftms_sub))

```

Visualizing spectra data
```{r}
sp <- spectra(ftms_sub)
spectraData(sp)
```


```{r}
cps <- chromPeakSpectra(ftms_sub, ppm = 10)
cps$precursorMz
```

Visualizing some peaks:
```{r}
peaksData(sp)[[2]]
```

```{r}
peaksData(sp)[[126]]
```
#Associate MS2 spectra to features

We remove the MS1 spectra, and keep juste the MS2, MS3, and MS4 spectra:
```{r}
sps <- spectra(ftms_sub)

frag_sps <-  chromPeakSpectra(ftms_sub, msLevel = 2, ppm = 10)

#spectraData(frag_sps)$feature_id <- "FT001"

spectraData(frag_sps)

msLevel(frag_sps) |>
  table()
```
Rename the chrom_peak_id column to feature_id for more compatibility with 
the workflow
```{r}
frag_sps$feature_id <- frag_sps$chrom_peak_id
frag_sps$feature_mzmed <- frag_sps$chrom_peak_mz
frag_sps$feature_rtmed <- frag_sps$chrom_peak_rt
```

#Grouping MSn spectra into spectral trees

We next extract the full MSn tree for the feature:
```{r}
ftms_msn_tree <- ftms_all_levels(ftms,frag_sps)
```
Asigning an id for each tree
```{r}
ftms_msn_tree_herba <- assign_msntree_id(ftms_msn_tree)
spectraVariables(ftms_msn_tree_herba)
```
```{r}
length(unique(ftms_msn_tree_herba$MSntreeID))
ftms_msn_tree_herba$MSntreeID |>
    table() |>
    quantile()

table(table(ftms_msn_tree_herba$MSntreeID))
```
#Select single best MSn spectral tree per feature

Select the feature with the highest basePeakIntensity
```{r}
## Metadata
spd <- spectraData(ftms_msn_tree_herba)

## Find the MS2 spectrum with the highest base peak intensity
ms2 <- spd$msLevel == 2

best_tree <- spd$MSntreeID[
  which.max(spd$basePeakIntensity[ms2])
]

## Keep the complete tree
best_ft <- ftms_msn_tree_herba[
  spd$MSntreeID == best_tree
]
spectraData(best_ft)
```

Enrich the tree
```{r}
frag_dt <- as.data.frame(spectraData(ftms_msn_tree_herba))
ftms_final <- enrich_msntree(
  ftms_best = best_ft,
  sd_filtered = frag_dt
)
```

```{r}
spectraData(ftms_final)
```
filter out all peaks that have intensity below 1% (or 0.5%) of the base peak

```{r}
filter_basepeak_intensity <- function(x, threshold = 0.01) {
  
  if (nrow(x) == 0)
    return(x)
  
  base_peak <- max(x[, 2], na.rm = TRUE)
  
  x[x[, 2] >= threshold * base_peak, , drop = FALSE]
}
```

```{r}
ftms_msn_tree_best <- addProcessing(
  ftms_final,
  function(x) {
    filter_basepeak_intensity(x, threshold = 0.01)
  }
)
```

```{r}
cbind(
  before = lengths(peaksData(ftms_final)),
  after  = lengths(peaksData(ftms_msn_tree_best))
)
```


Now the spectra object contains just one tree. 

Save the object with trees localy:
```{r}
save(ftms_msn_tree_best, file = here("data", "ftms_msn_tree_best.RData"))
```

#Pseudo MS2 spectra

Create assembled tree (pseudo MS2) by calling the assembledSpectraTree() function
from RDynlib package:

```{r}

ftms_assembled_herba <- msntree_to_pseudoMS2(ftms_msn_tree_best)

ftms_assembled_herba
unique(spectraData(ftms_assembled_herba)$msLevel)
```

```{r}
spectraData(ftms_assembled_herba)
```
Add "not_assembled" label to the initial object with separated spectrums 
```{r}
spectraData(ftms_msn_tree_best)$spectrum.type <- "individual"
```

Concatenate both spectra objects with "pseudo_MS2" and "individual" into one
spectra object.
```{r}
ftms_one_merged_tree <- setBackend(ftms_msn_tree_best,  MsBackendMemory())
ftms_assembled_herba <- setBackend(ftms_assembled_herba,  MsBackendMemory())

# Concatenate
ftms_all_herba <- concatenateSpectra(ftms_one_merged_tree, ftms_assembled_herba)

peaksData(ftms_all_herba)[[1,]]            

head(spectraData(ftms_all_herba))
```
save locally the object that contains both spectrum types
```{r}
save(ftms_all_herba, file = here("data", "ftms_all_herba.RData"))
```


#Create sql database for this feature
```{r}
load("data/ftms_all_herba.RData")
dbfile <- here("data", "herba_ftneg.sqlite")
createSpectraSQLite(
    sps = ftms_all_herba,
    dbfile = dbfile,
    date = '01/06/2015',
    user = 'ament',
    machine = 'LTQ_Orbitrap',
    column = 'Waters_Sunfire_C18',
    mstype = 'FT',
    buffera = 'water_AA',
    bufferb = 'ACN',
    gradient_time = '65(95-95in2_to82at32_to76at60min__to0at65minat300_40gr) min',
    source = 'ESI',
    mode = 'neg',
    species = 'Linum usitatissimum',
    tissue = 'seed_integument',
    ce = '35',
    meta = 'Linea_varieties_S1_S9'
)
```


# Add the herbacytine feature to Dynlib sql
```{r}
dbf0 <- here("data","DYL_FLAX_FTMSneg.sqlite")
db <- dbConnect(SQLite(), dbf0)
```

retreive the maximum expid in the global database before adding the new feature.
```{r}
dbGetQuery(db, "
  SELECT MAX(CAST(expid AS INTEGER)) AS max_expid
  FROM experiment
")
```
#Add the new experiment to the existing Dynlib
```{r}
# Create new database
#dbf <- here("data", "DYL_herba2_FTMSneg.sqlite")
#con_merged <- dbConnect(RSQLite::SQLite(), dbf)

# merge
merge_databases(
  main_db   = here("data","DYL_FLAX_FTMSneg.sqlite"),
  add_db    = here("data","herba_ftneg.sqlite"),
  output_db = here("data","DYL_flax_herba_FTMSneg.sqlite")
)
```
## connect to the merged database
```{r}
dbf <- here("data", "DYL_flax_herba_FTMSneg.sqlite")
con_merged <- dbConnect(RSQLite::SQLite(), dbf)
```

Retreive the new maximum expid (it should be the old expid + 1)
```{r}
dbGetQuery(con_merged, "
  SELECT MAX(CAST(expid AS INTEGER)) AS max_expid
  FROM experiment
")
```

```{r}
add_db <- here("data","herba_ftneg.sqlite")
con_herba <- dbConnect(RSQLite::SQLite(), add_db)
dbGetQuery(con_herba, "Select * from experiment ")
```

```{r}
dbExecute(
    con_merged,
    "
    UPDATE experiment
    SET
        date = '01/06/2015',
        user = 'ament',
        machine = 'LTQ_Orbitrap',
        column = 'Waters_Sunfire_C18',
        mstype = 'FT',
        buffera = 'water_AA',
        bufferb = 'ACN',
        gradient_time = '65(95-95in2_to82at32_to76at60min__to0at65minat300_40gr) min',
        source = 'ESI',
        mode = 'neg',
        species = 'Linum usitatissimum',
        tissue = 'seed_integument',
        ce = '35',
        meta = 'Linea_varieties_S1_S9'
    WHERE expid = 266
    "
)
```


```{r}
dbExecute(
    con_merged,
    "
    UPDATE experiment
    SET
        date = '01/06/2015',
        user = 'rebdau',
        machine = 'LTQ_Orbitrap',
        column = 'Waters_Sunfire_C18',
        mstype = 'FT',
        buffera = 'water_AA',
        bufferb = 'ACN',
        gradient_time = '65(95-95in2_to82at32_to76at60min__to0at65minat300_40gr) min',
        source = 'ESI',
        mode = 'neg',
        species = 'Standards',
        tissue = 'seed_integument',
        ce = '35',
        meta = 'Linea_varieties_S1_S9'
    WHERE expid = 267
    "
)
```


```{r}
dbGetQuery(con_merged, "Select * from experiment where expid = '266' or expid = '267' ")
```

Fill the herbacetin data by the EnterName() function:

```{r}
EnterName_SQL(dbf, 101307)
```

```{r}
dbGetQuery(con_merged, "Select * from ms_compound where expid = '267' ")
``` 

Change the "spectrum_type" of the existing dynlib from "NA" to "individual":
```{r}
DBI::dbExecute(
  con_merged,
  "UPDATE msms_spectrum
   SET spectrum_type = 'individual'
   WHERE spectrum_type IS NULL"
)
```

Verification:
```{r}
DBI::dbGetQuery(
  con_merged,
  "SELECT spectrum_type FROM msms_spectrum"
)
```
Convert the sql database to spectra object:
```{r}
dbf <- here("data", "DYL_flax_herba_FTMSneg.sqlite")
dy_sps <- sql_to_spectra(dbf)
```

Add accessions to the database with specefic initials for flax and herbacetin 
experiments, the other experiments would have the first two letters of the user
column from experiment table as initials.
```{r}
add_accessions(dbf, 
                initials_overrides = c("266" = "RD",
                                    "267" = "RD"))
```


```{r}
MSnplot_SQL(dbf, 101307)
```


```{r}
MSnplot_SQL(dbf, 101280)
```

