---
title: "cspp and gnps-like networks of known molecules"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

In this quarto document we plot the cspp and gnps-like networks of known 
molecules from the aligned compounds of ftms and qtof flax data. 


```{r}
#| message: false

#' Load required libraries
library(devtools)
library(DBI)
library(RSQLite)
library(here)
library(data.table)
#rm(list = c("convert_compid_assoc"))
#` Uploading RDynLib: Load all functions of the RDynlib directory
load_all()
```

#FTMS flax negative

```{r}
ft_path <- here("data", "DYL_FLAX_FTMSneg.sqlite")
con <- dbConnect(RSQLite::SQLite(), ft_path)
cspp_ft <- here("data","cspp_flax.txt")
charge_file <- here("data", "ftms_all_compounds_RD.xlsx")
```

Add additional columns to compound_add table:
```{r}
library(DBI)
library(RSQLite)
library(here)

ft_path <- here("data", "DYL_FLAX_FTMSneg.sqlite")
con <- dbConnect(RSQLite::SQLite(), ft_path)

# Add columns
dbExecute(con, "ALTER TABLE compound_add ADD COLUMN Suc TEXT")
dbExecute(con, "ALTER TABLE compound_add ADD COLUMN HMG TEXT")
dbExecute(con, "ALTER TABLE compound_add ADD COLUMN Her TEXT")
dbExecute(con, "ALTER TABLE compound_add ADD COLUMN Hdg TEXT")
dbExecute(con, "ALTER TABLE compound_add ADD COLUMN NiA TEXT")
dbExecute(con, "ALTER TABLE compound_add ADD COLUMN Hmg_C TEXT")

# Optional renamed version instead of Hmg
dbExecute(con, "ALTER TABLE compound_add DROP COLUMN V36")
dbExecute(con, "ALTER TABLE compound_add DROP COLUMN V37")
dbExecute(con, "ALTER TABLE compound_add DROP COLUMN V38")
dbExecute(con, "ALTER TABLE compound_add DROP COLUMN V39")
dbExecute(con, "ALTER TABLE compound_add DROP COLUMN V40")

```

cspp conversions
```{r}
#cspp_ft_2ch <- here("data","cspp_flax_2ch.txt")
cspp.tot_SQL(ft_path, 266, mzerr = 0.01,
                         cspp = cspp_ft, peakwidth = 0.1, IntThres  = 0, mz_tol = 0.5,
                         dot_thresh = 0, spectrum_type = "assembled",  
                         min_common = 0, charge_file = charge_file)
```


```{r}
dbGetQuery(con, "select * from compound_add where compound_id = 101280")
```
```{r}
targMS2comp_SQL(101099, 101280, ms2_list)
```



```{r}
dbGetQuery(con, "
  SELECT compound_id, HEX
  FROM compound_add
  WHERE HEX IS NOT NULL
    AND HEX != 'NULL'
  LIMIT 10
")
```


Test if there is some new cspp conversions
```{r}
db_ft <- dbConnect(SQLite(), ft_path)
DBI::dbGetQuery(db_ft, "
  SELECT g.*
  FROM compound_add g
  INNER JOIN ms_compound c 
      ON g.compound_id = c.compound_id
  WHERE c.expid = 266 
    AND (
         g.GNPS IS NOT NULL OR
         g.RED IS NOT NULL OR
         g.ROX IS NOT NULL OR
         g.MET IS NOT NULL OR
         g.OXY IS NOT NULL OR
         g.HYD IS NOT NULL OR
         g.BOX IS NOT NULL OR
         g.CSPP_MOX IS NOT NULL OR
         g.CSPP_ACE IS NOT NULL OR
         g.CSPP_CAR IS NOT NULL OR
         g.CSPP_PRE IS NOT NULL OR
         g.CSPP_GLY IS NOT NULL OR
         g.CSPP_MON IS NOT NULL OR
         g.CSPP_HdN IS NOT NULL OR
         g.CSPP_HBA IS NOT NULL OR
         g.CSPP_PEN IS NOT NULL OR
         g.CSPP_COU IS NOT NULL OR
         g.CSPP_RHA IS NOT NULL OR
         g.CSPP_HuN_C IS NOT NULL OR
         g.CSPP_VAN IS NOT NULL OR
         g.CSPP_HaN_C IS NOT NULL OR
         g.CSPP_HEX IS NOT NULL OR
         g.CSPP_DCA IS NOT NULL OR
         g.CSPP_HUN IS NOT NULL OR
         g.CSPP_URO IS NOT NULL OR
         g.CSPP_FER IS NOT NULL OR
         g.CSPP_GuN_C IS NOT NULL OR
         g.CSPP_SYR IS NOT NULL OR
         g.CSPP_GaN_C IS NOT NULL OR
         g.CSPP_GUN IS NOT NULL OR
         g.CSPP_SIN IS NOT NULL OR
         g.CSPP_SuN_C IS NOT NULL OR
         g.CSPP_SaN_C IS NOT NULL OR
         g.CSPP_SUN IS NOT NULL
        )
  ORDER BY g.compound_id
")
```

#GNPS conversion
```{r}
conv.GNPS_SQL(sql_path= ft_path, expid = 266,
                          peakwidth = 0.2,
                          mzerr = 0.015,
                          min = 5,
                          adduct = 60.0211, #what should I put here?
                          thr1 = 2,
                          thr2 = 0.1,
                          thr3 = 0.3,
                          spectrum_type = "not_assembled", 
                          IntThres = 0)
```


```{r}
db_ft <- dbConnect(SQLite(), ft_path)
DBI::dbGetQuery(db_ft, "
  SELECT g.*
  FROM gnps_add g
  INNER JOIN ms_compound c 
      ON g.compound_id = c.compound_id
  WHERE c.expid = 266 
    AND (
         g.GNPS_1 IS NOT NULL OR
         g.GNPS_2 IS NOT NULL OR
         g.GNPS_3 IS NOT NULL OR
         g.GNPS_4 IS NOT NULL OR
         g.GNPS_5 IS NOT NULL 
        )
  ORDER BY g.compound_id
")
```


Visualize the distribustion of the similarity scores
```{r}
quantile(edges.raw$ave.common, c(0.25, 0.5, 0.75))
quantile(edges.raw$ave.dot, c(0.25, 0.5, 0.75))
thr2 <- quantile(edges.raw$ave.common, 0.25)
thr3 <- quantile(edges.raw$ave.dot, 0.25)
```

Plot the network of the HDG-succinate-HDG with a mass of 666,14:
```{r}
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 100988, min = 0,
                            nr_of_seq = 2,
                           thr1 = 1,
                          thr2 = 0.3, 
                          thr3 = 0.3)
```


Plot the network of the HDG-succinate-HDG with a mass of 666,14:
```{r}
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 101283, min = 0,
                            nr_of_seq = 2,
                           thr1 = 1,
                          thr2 = 0.3, 
                          thr3 = 0.3)
```

Plot the network of the HDG-succinate-HDG with a mass of 747,16:
```{r}
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 101294, min = 0,
                            nr_of_seq = 2,
                           thr1 = 1,
                          thr2 = 0.3, 
                          thr3 = 0.3)
```

Plot the network of the HDG-succinate-HDG with a mass of 747,16:
```{r}
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 101213, min = 0,
                            nr_of_seq = 2,
                           thr1 = 1,
                          thr2 = 0.3, 
                          thr3 = 0.3)
```

```{r}
load_all()
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 101280, min = 0,
                            nr_of_seq = 3,
                           thr1 = 0.1,
                          thr2 = 0.2, 
                          thr3 = 0.3)
```

```{r}
load_all()
addit <- net.addit_SQL(sql_path = ft_path, exp.id = 266, nettype = "CSPP",
                          nr_col = NULL,
                          min = 100,
                          thr1 = 2,
                          thr2 = 0.3,
                          thr3 = 0.3)

Local.net_SQL(
  net.lst = addit,
  dbkey      = "101280",
  nettype    = "CSPP",
  output_png = "network_CSPP2.png",   # omit to just display in RStudio
  png_width  = 3000,
  png_height = 3000,
  png_res    = 150
)
```




```{r}
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 101007, min = 0,
                            nr_of_seq = 2,
                           thr1 = 0.1,
                          thr2 = 0.3, 
                          thr3 = 0.3)
```


#Flax Qtof negtive

First get the corresponding aligned compounds of qtof data based on ftms compounds.
```{r}
#loading the txt file that contains the alignment results
alignment_file <- here("data", "assoc_flax_63.txt")

getAlignedCompids <- getAlignedCompids_SQL(
compound_id = c(
  100926, 100998, 100974, 100934, 101238, 100946, 101211, 100988, 101242, 
  101097, 101141, 101019, 101147, 100991, 101098, 101296, 101154
),
  database_name = "DYL_FLAX_FTMSneg.sqlite",
  alignment_file = alignment_file
)
print(getAlignedCompids)
```



```{r}
conv.GNPS_SQL(sql_path= qt_path, expid = 52,
                          mzerr = 0.015,
                          min = 5,
                          adduct = 60.0211,
                          thr1 = 3,
                          thr2 = 0.1,
                          thr3 = 0.4,
                          spectrum_type = "intersect_single_energy", 
                          IntThres = 0)
```

```{r}
db_qt <- dbConnect(SQLite(), qt_path)
DBI::dbGetQuery(db_qt, "
  SELECT g.*
  FROM gnps_add g
  INNER JOIN ms_compound c 
      ON g.compound_id = c.compound_id
  WHERE c.expid = 52 
    AND (
         g.GNPS_1 IS NOT NULL OR
         g.GNPS_2 IS NOT NULL OR
         g.GNPS_3 IS NOT NULL OR
         g.GNPS_4 IS NOT NULL OR
         g.GNPS_5 IS NOT NULL 
        )
  ORDER BY g.compound_id
")
```

#cspp conversion
```{r}
cspp.tot_SQL(qt_path, 52, mzerr = 0.01,
                         cspp = cspp_ft, peakwidth = 0.2, IntThres  = 0, mz_tol = 0.5,
                         dot_thresh = 0.7, spectrum_type = "intersect_single_energy",  
                         min_common = 3, charge_file = charge_file)
```



plot ntworks
```{r}
Overall.net_SQL(sql_path = qt_path, exp.id = 52, dbkey = 45843, min = 0,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```


```{r}
Overall.net_SQL(sql_path = qt_path, exp.id = 52, dbkey = 45858, min = 0,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```


```{r}
Overall.net_SQL(sql_path = qt_path, exp.id = 52, dbkey = 45847, min = 0,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```

```{r}
Overall.net_SQL(sql_path = qt_path, exp.id = 52, dbkey = 45890, min = 0,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```

```{r}
Overall.net_SQL(sql_path = qt_path, exp.id = 52, dbkey = 45918, min = 0,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```


```{r}
Overall.net_SQL(sql_path = qt_path, exp.id = 52, dbkey = 45848, min = 0,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```



