---
title: "workflow of adding names of  newly characterized Compounds"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This quarto document represents the workflow for calling the `NodenameMass_SQL()` 
to show all the ms_compoud table of a given compound based on the mass, 
retention time, name or node_name. then using the returned compound_id we can 
call the `EnterName_SQL()` function to Interactively annotate and update 
compound names in an SQLite database.



# Required packages

This project follows an R package layout as follows:

The main project folder is:

RDynLib_SQL/

Within this directory, the structure is organized as follows:

RDynLib_SQL/ ├── data/ ├── R/ └── vignettes/

### `data/`

Contains the SQLite database files used in the workflow (e.g., `ftms_flax_neg.sqlite`).

### `R/`

Contains the R source files (`.R`) that define the package functions and core implementation.

### `vignettes/`

Contains Quarto (`.qmd`) files that document the workflow and provide reproducible examples.

In this workflow, the working directory is set to the project root:

RDynLib_SQL/

## Loading required packages

```{r}
library(here)
library(Spectra)
library(DBI)
library(RSQLite)
library(devtools)

#` Uploading RDynLibSQL: Load all functions of the RDynlib directory
load_all()
```

#(1) database file: 

```{r}
dbf <- here("data", "DYL_FLAX_FTMSneg.sqlite")
db <- dbConnect(SQLite(), dbf)
```


#(2) Print the ms_compound table of a given masses and retention time range:
```{r}
NodenameMass_SQL(dbf, expid = 266,
                             mass_measured = 1083.28,
                             rt = 26.76,
                             mass_range = 0,
                             rt_range = 0)
```


#(3) Update the ms_compound fields related to the previously extracted
# compound_id using the NodenameMass_SQL: 
```{r}
EnterName_SQL(dbf, CoMPID = )
```

