---
title: "Create an sql subdatabase of a given experiment id"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

Quarto document to demonstrate how to use filterDynLibDB() function from RDynlib 
package to create an sql subdatabase of a given experiment

```{r}
#| message: false
#| results: "hide"
#| warning: false
#' Load required libraries
library(devtools)
library(DBI)
library(RSQLite)
library(here)
#rm(list = c("convert_compid_assoc"))
#` Uploading RDynLib: Load all functions of the RDynlib directory
load_all()
```

#File paths to the input and output sql databases
```{r}
input_db <- here("data", "DYL_FLAX_FTMSneg.sqlite")
output_db <- here("data", "filter_flax.sqlite")
output_db2 <- here("data", "filter_5_7_12.sqlite")

```

#Call the function to filter the database based on a single expid
```{r}
filterDynLibDB(
  expid = 266,
  input_db = input_db,
  output_db = output_db
)
```

#Connection to the output database
```{r}
con_filter <- dbConnect(RSQLite::SQLite(), output_db)

#Verify the experiment table
dbGetQuery(con_filter , "
SELECT *
FROM experiment
")
```

#Print ms_compound table of the filtered database
```{r}
dbGetQuery(con_filter , "
SELECT *
FROM ms_compound
")
```


#Call the function to filter the database based on a set of expids
```{r}
filterDynLibDB(
    expid = c(5, 7, 12),
    input_db = input_db,
    output_db = output_db2
)
```



#Connection to the output2 database
```{r}
con_filter2 <- dbConnect(RSQLite::SQLite(), output_db2)

#Verify the experiment table
dbGetQuery(con_filter2 , "
SELECT *
FROM experiment
")
```