---
title: "Create an sql subdatabase of a given experiment id"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

Quarto document to demonstrate how to use filterDynLibDB() function from RDynlib 
package to create an sql subdatabase of a given experiment

```{r}
#| message: false
#| results: "hide"
#| warning: false
#' Load required libraries
library(devtools)
library(DBI)
library(RSQLite)
library(here)
library(RDynLib)
#rm(list = c("convert_compid_assoc"))
#` Uploading RDynLib: Load all functions of the RDynlib directory
load_all()
```

#File paths to the input and output sql databasesftms
```{r}
input_db <- here("data", "DYL_flax_herba_FTMSneg.sqlite")
output_db <- here("data", "Ftmsflax_266_267.sqlite")
```

#Call the function to filter the database based on a single expid
```{r}
filterDynLibDB(
  expid = c(266,267),
  input_db = input_db,
  output_db = output_db
)
```

#Connection to the output database
```{r}
con_filter <- dbConnect(RSQLite::SQLite(), output_db)

#Verify the experiment table
dbGetQuery(con_filter , "
SELECT *
FROM experiment
")
```

#Print ms_compound table of the filtered database
```{r}
dbGetQuery(con_filter , "
SELECT *
FROM ms_compound
")
```
```{r}
dbGetQuery(con_filter , "
SELECT *
FROM msms_spectrum
")
```


```{r}
dbGetQuery(con_filter, "
SELECT COUNT(*) AS n_rows
FROM ms_compound
WHERE expid = 266  
")
```
```{r}
dbGetQuery(con_filter, "
SELECT COUNT(*) AS n_rows
FROM msms_spectrum s
JOIN ms_compound c
  ON s.compound_id = c.compound_id
WHERE c.expid = 266 and s.ms_level = 4 and s.spectrum_type = 'not_assembled'
")
```
```{r}
dbGetQuery(con_filter, "
SELECT COUNT(*) AS n_rows
FROM msms_spectrum_peak p
JOIN msms_spectrum s
  ON p.spectrum_id = s.spectrum_id
JOIN ms_compound c
  ON s.compound_id = c.compound_id
WHERE c.expid = 267
")
```

#QTOF
#File paths to the input and output sql databases
```{r}
input_qt <- here("data", "DyLib_QTOF_pos.sqlite")
output_qt <- here("data/database_to_publish", "qtof_52.sqlite")
```

#Call the function to filter the database based on a single expid
```{r}
filterDynLibDB(
  expid = 52,
  input_db = input_qt,
  output_db = output_qt
)
```

#Connection to the output database
```{r}
qt_filter <- dbConnect(RSQLite::SQLite(), output_qt)

#Verify the experiment table
dbGetQuery(qt_filter , "
SELECT *
FROM msms_spectrum 
")
```

```{r}
dbGetQuery(qt_filter, "
SELECT COUNT(*) AS n_rows
FROM feature_matrix
WHERE expid = 52  
")
```
```{r}
dbGetQuery(qt_filter, "
SELECT COUNT(*) AS n_rows
FROM msms_spectrum s
JOIN ms_compound c
  ON s.compound_id = c.compound_id
WHERE c.expid = 52 and s.ms_level = 2
")
```
```{r}
dbGetQuery(qt_filter, "
SELECT COUNT(*) AS n_rows
FROM msms_spectrum
WHERE spectrum_type = 'intersect_single_energy'
  AND dataOrigin LIKE '%DDA1%'
")
```

```{r}
dbGetQuery(qt_filter, "
SELECT COUNT(*) AS n_rows
FROM msms_spectrum
WHERE spectrum_type = 'merged_MSMS_all_energies'
")
```

