---
title: "Adding flax ftms neg sql to Dynlib ftms neg sql"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
editor: 
  markdown: 
    wrap: 72
---

## Shiny Documents

This quarto document represents the full workflow to add ftmsneg flax to
Dynlib sql database, and then to a CompDb and spectra Objects. The
retention times in Dynlib are in minutes, but in flax data are in
seconds. In the following workflow while adding flax to Dynlib we will
unify the units and convert the retention times of flax data to minutes.

Note: In this file we will need the similarity output file between the
existing database and the new experiment that we would like to  add,
that's why we have to run first this quarto document: \
[FTMS_neg_call.qmd](https://github.com/rebdau/RDynLib/blob/dev_Ahlam/Similarity/FTMS/FTMS_neg_call.qmd)
and load the output here.

# Directroy layout

This project follows an  R package layout as follows:

The main project folder is:

RDynLib_SQL/


Within this directory, the structure is organized as follows:

RDynLib_SQL/
├── data/
├── R/
└── vignettes/


### `data/`

Contains the SQLite database files used in the workflow (e.g., `ftms_flax_neg.sqlite`).

### `R/`

Contains the R source files (`.R`) that define the package functions and core implementation.

### `vignettes/`

Contains Quarto (`.qmd`) files that document the workflow and provide reproducible examples.

In this workflow, the working directory is set to the project root:

RDynLib_SQL/

# Required packages

```{r}
library(here)
library(Spectra)
library(DBI)
library(dplyr)
library(RSQLite)
library(CompoundDb)
library(devtools)

#` Uploading RDynLibSQL: Load all functions of the RDynlib directory
load_all()

register(SerialParam())

```

# Load databases
Loading the sql database of flax ftms negative

```{r}
dbf <- here("data", "ftms_flax_neg.sqlite")
con_flax <- dbConnect(RSQLite::SQLite(), dbf)
```

Loading the sql database of Dynlib qtof negative

```{r}
dbf <- here("data", "DyLib_FTMS_neg_or.sqlite")
con_dyL <- dbConnect(RSQLite::SQLite(), dbf)
```

```{r}
dbListTables(con_flax)
```

```{r}
dbListTables(con_dyL)
```

# Merge databases
```{r}
# Create new database
dbf <- here("data", "DYL_FLAX_FTMSneg.sqlite")
con_merged <- dbConnect(RSQLite::SQLite(), dbf)

# merge
merge_databases(
  main_db   = here("data","DyLib_FTMS_neg_or.sqlite"),
  add_db    = here("data","ftms_flax_neg.sqlite"),
  output_db = here("data","DYL_FLAX_FTMSneg.sqlite")
)


```

# Check merging
```{r}
dbListTables(con_merged)
```

```{r}
dbGetQuery(con_merged, "PRAGMA table_info(msms_spectrum);")
dbGetQuery(con_merged, "PRAGMA table_info(ms_compound);")
```

### Ensuring the experiment table has been well merged

```{r}
dbReadTable(con_merged, "experiment")
```

```{r}
dbGetQuery(con_merged, "Select * from experiment where user = 'Ahlam'")
```

The new experiment has 266 as expid

### Ensuring the ms_compound has been well merged

```{r}
dbGetQuery(con_dyL, "
  SELECT * FROM ms_compound
  ORDER BY CAST(compound_id AS INTEGER) DESC
  LIMIT 10
")

```

The highest compound_id is 100911 in the Dynlib database

```{r}
dbGetQuery(con_merged, "Select * from ms_compound where expid = 266 LIMIT 10")
```

the query results show that the compound_id of the added experiment
"266" start by 100912 which is correct

### Ensuring the msms_spectrum table has been well merged

```{r}
dbGetQuery(con_dyL, "
  SELECT * FROM msms_spectrum
  ORDER BY CAST(spectrum_id AS INTEGER) DESC
  LIMIT 10
")
```

The highest spectrum_id is 201034 in the Dynlib database

```{r}
dbGetQuery(con_merged, "Select * from msms_spectrum where compound_id = 100912 Limit 10")
```

the query results show that the spectrum_id of the added experiment
"266" start by 201035 which is correct

### Ensuring the msms_spectrum_peak table has been well merged

```{r}
dbGetQuery(con_merged, "Select * from msms_spectrum_peak where spectrum_id = 201035")
```

### Ensuring the feature_matrix table has been well merged

```{r}
dbReadTable(con_merged, "feature_matrix")
```

```{r}
dbGetQuery(con_merged, "Select * from feature_matrix where expid = 266 LIMIT 10")
```

# Converting the merged sql connection to a compdb object

```{r}
cdb <- CompDb(con_merged)
cdb
```

```{r}
compounds(cdb) |> head()
```

```{r}
sp <- Spectra(cdb)
```

```{r}
spectraData(sp, columns = c("msLevel", "compound_id", "nodename", "rtime", "retention_time", "spectrum_type"))
```

# Load the similarity table between flax and Dynlib ftms negative data
We need the similarity output between the
existing database and the new experiment that we added. The object was called `ftms_matches`, and generated
in this quarto document: \
[FTMS_neg_call.qmd](https://github.com/rebdau/RDynLib/blob/dev_Ahlam/Similarity/FTMS/FTMS_neg_call.qmd).
We load the object here.

```{r}
load(here("data","ftms_matches.RData"))
print(ftms_matches)
```


# Add automatically generated names of matching compounds to the merged experiment in compoundDb

connect to the database:
```{r}
dbf <- here("data", "DYL_FLAX_FTMSneg.sqlite")
con_merged <- dbConnect(RSQLite::SQLite(), dbf)
```


```{r}
adding_Names_SQL(con_merged, ftms_matches)
```

```{r}
dbGetQuery(con_merged, "Select * from ms_compound where expid = 266 and name IS NOT NULL")
```

# Disonnect databases
```{r}
dbDisconnect(con_flax)
dbDisconnect(con_dyL)
dbDisconnect(con_merged)
```

