---
title: "workflow to test the similarity_RDynlib() function on ftms data"
format: html
  server: shiny
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This quarto document represents the workflow for calling the `similarity_RDynlib()` from the [similarity_functions.R](https://github.com/rebdau/RDynLib_SQL/blob/Ahlam/R/similarity_functions.R) file to calculate the similarity between FTMS flax data and the RDynLib FTMS database.

# Required packages

This project follows an R package layout as follows:

The main project folder is:

RDynLib_SQL/

Within this directory, the structure is organized as follows:

RDynLib_SQL/ ├── data/ ├── R/ └── vignettes/

### `data/`

Contains the SQLite database files used in the workflow (e.g., `ftms_flax_neg.sqlite`).

### `R/`

Contains the R source files (`.R`) that define the package functions and core implementation.

### `vignettes/`

Contains Quarto (`.qmd`) files that document the workflow and provide reproducible examples.

In this workflow, the working directory is set to the project root:

RDynLib_SQL/

## Loading required packages

First, install the new CompoundDb version:

```{r}
BiocManager::install("RforMassSpectrometry/CompoundDb", ref = "jomain", force = TRUE)
```

```{r}
library(here)
library(Spectra)
library(CompoundDb)
library(MetaboAnnotation)
library(DBI)
library(RSQLite)
library(devtools)

#` Uploading RDynLibSQL: Load all functions of the RDynlib directory
load_all()
```

## convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

Next we convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

```{r}

dbf <- here("data", "DyLib_FTMS_neg_or.sqlite")
db <- dbConnect(SQLite(), dbf)
dbListTables(db)
#dbExecute(db, "DROP TABLE IF EXISTS feature_matrix ")
experiment_df <- dbGetQuery(db, "SELECT * FROM experiment")

#here we remove "x", "x.1" columns from experiment table 
experiment_df <- experiment_df[, !names(experiment_df) %in% c("x", "x.1")]
dbExecute(db, "DROP TABLE experiment")
dbWriteTable(db, "experiment", experiment_df)

#the columns have been successfully deleted
dbGetQuery(db, "PRAGMA table_info(experiment)")
dbDisconnect(db)
#we convert the sql database to a Compdb object
cdb <- ExpCompDb(dbf)
#we convert the Compdb object to a spectra object
dy_sps <- Spectra(cdb)
```

```{r}
spectraData(dy_sps, columns = c("retention_time"))
```

Here we load the FTMS of flax seed database that contains both spectrum types : the assembled and not assembled one (it's already a spectra object),\
you can load it from: [ftmsneg_flax_sql.qmd](https://github.com/rebdau/RDynLib_ftms/blob/ahlam/ftmsneg_flax_sql.qmd)

```{r}

load(here("data","database_FTMS_neg","ftms_all.RData"))
ftms_all <- ftms_all
spectraData(ftms_all)

```

Next we calculate the similarity score between the dynlib and flax seed FTMS negative integuments.

#test the new function(where the user should enter the spectrum type and mslevel )

```{r}

ftms_matches <- similarity_RDynlib_match(
  ftms_all,
  dy_sps,
  polarity_query = 0,
  polarity_target = 0,
  requirePrecursor = TRUE,
  tolerance = 0,
  ppm = 5,
  fragments_resolution = "unit.resolution",
)


```

print the similarity matrix and the number of rows (matches)

```{r}

print(ftms_matches)
nrow(ftms_matches)
```

save the resulting object:

```{r}
save(ftms_matches, file = here("data", "ftms_matches.RData"))

```

```{r}
load(here("data","ftms_matches.RData"))
```

