---
title: "Untitled"
format: html
server: shiny
---

In this quarto document we call the "sim_functions.R" file code to use functions for similarity calculation between FTMS data and the RDynLib FTMS database.

```{r}
library(Spectra)
library(CompoundDb)
pth <- getwd()
pth
setwd("C:/Users/ament/Desktop/Github/RDynLib/Similarity/FTMS")
pth <- getwd()
pth
```

First we load the "sim_functions.R" file that contains functions for similarity calculation.

```{r}
source("sim_functions_ftmsneg.R")
```

Here we install the new CompoundDb version

```{r}
BiocManager::install("RforMassSpectrometry/CompoundDb", ref = "jomain", force = TRUE)
```

```{r}
ExpCompDb <- function(x) {
    cdb <- CompDb(x)
    CompoundDb::addJoinDefinition(cdb, "ms_compound", "experiment", "expid", "expid")
}
```

Next we convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

```{r}
library(DBI)
library(RSQLite)
dbf <- file.path(pth,"/data", "DyLib_FTMS_neg_or.sqlite")
db <- dbConnect(SQLite(), dbf)
dbExecute(db, "DROP TABLE IF EXISTS feature_matrix ")
experiment_df <- dbGetQuery(db, "SELECT * FROM experiment")

#here we remove "x", "x.1" columns from experiment table 
experiment_df <- experiment_df[, !names(experiment_df) %in% c("x", "x.1")]
dbExecute(db, "DROP TABLE experiment")
dbWriteTable(db, "experiment", experiment_df)

#the columns have been successfully deleted
dbGetQuery(db, "PRAGMA table_info(experiment)")
dbDisconnect(db)
#we convert the sql database to a Compdb object
cdb <- ExpCompDb(dbf)
#we convert the Compdb object to a spectra object
dy_sps <- Spectra(cdb)
```

```{r}
spectraData(dy_sps, columns = c("rtime"))
```

Here we load the FTMS of flax seed database (it's already a spectra object)\
you can load it from: [ftmsneg_flax_sql.qmd](https://github.com/rebdau/RDynLib_ftms/blob/ahlam/ftmsneg_flax_sql.qmd)

```{r}
load("D:/work/these/These/clones/new_clones/RDynLib/Similarity/FTMS/data/ftms_all.RData")
ftms_all <- ftms_all
spectraData(ftms_all)
```

Next we calculate the similarity score between the dynlib and flax seed FTMS negative integuments.

#test the new function(where the user should enter the mslevel and spectrum type)

```{r}
source("D:/work/these/These/clones/new_clones/RDynLib/Similarity/FTMS/sim_new.R")

res1 <- symmetric_dotproduct_combined(
  ftms_all,
  dy_sps,
  polarity_query = 0,
  polarity_target = 0,

  precursor_resolution = "unit.resolution",
  fragments_resolution = "unit.resolution",
)


```
