---
title: "compound_id of identified compounds"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

In this Quarto document we extract the compound_id from flax qtof negative that 
were aligned with the known compounds of flax ftms negative, and we use the 
MultMatchCID_Spectra() function to extract the matches of the flax ftms compounds, 
that were aligned with the qtof flax data, with the dynlib ftms database.


```{r}
#| message: false
#' Load required libraries
library(here)
library(Spectra)
library(CompoundDb)
library(MetaboAnnotation)
library(DBI)
library(RSQLite)
library(devtools)
library(openxlsx)
library(here)
#` Uploading RDynLib: Load all functions of the RDynlib directory
load_all()
```

Extracting the compounds that has been aligned with the following known ftms 
compounds using getAlignedCompids_SQL():
```{r}
#loading the txt file that contains the alignment results
alignment_file <- here("data", "assoc_flax.txt")

getAlignedCompids <- getAlignedCompids_SQL(
  compound_id = c(
    101294, 101283, 101280, 101242, 101238, 101216, 101214, 101213, 101211, 101155,
    101154, 101147, 101143, 101142, 101141, 101099, 101098, 101097, 101075, 101025,
    101019, 101010, 101007, 100991, 100988, 100977, 100951, 100946, 100934
  ),
  database_name = "DYL_FLAX_FTMSneg.sqlite",
  alignment_file = alignment_file
)
print(getAlignedCompids)
```


```{r}
save(getAlignedCompids, file = here("data", "getAlignedCompids.RData"))
```


## convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

Convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object to use it in 
MultmatchCID_spectra() function:

```{r}

dbf <- here("data", "DYL_FLAX_FTMSneg.sqlite")
db <- dbConnect(SQLite(), dbf)
dbListTables(db)
#dbExecute(db, "DROP TABLE IF EXISTS feature_matrix ")
experiment_df <- dbGetQuery(db, "SELECT * FROM experiment")

#here we remove "x", "x.1" columns from experiment table 
experiment_df <- experiment_df[, !names(experiment_df) %in% c("x", "x.1")]
dbExecute(db, "DROP TABLE experiment")
dbWriteTable(db, "experiment", experiment_df)

#the columns have been successfully deleted
dbGetQuery(db, "PRAGMA table_info(experiment)")
dbDisconnect(db)
#we convert the sql database to a Compdb object
cdb <- ExpCompDb(dbf)
#we convert the Compdb object to a spectra object
dy_sps <- Spectra(cdb)
```

```{r}
df <- spectraData(dy_sps, columns = c("msLevel", "compound_id", "expid", "retention_time"))
df[df$compound_id == "101242", ]
```

Next we run the MultMatchCID_Spectra() on the ftms compounds to extract their 
matches from the dynlib ftms database:

```{r}
#small test to the function
cmp <- MultMatchCID_Spectra(
  dy_sps,
  101280,
  spectrum_type = "not_assembled",
  ms_level_query = 2,
  ms_level_target = 2,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0,
  ppm = 10,
  minIons = 2,
  fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
cmp
```
```{r}
compound_ids <- c(
  101294, 101283, 101280, 101242, 101238, 101216, 101214, 101213, 101211, 101155,
  101154, 101147, 101143, 101142, 101141, 101099, 101098, 101097, 101075, 101025,
  101019, 101010, 101007, 100991, 100988, 100977, 100951, 100946, 100934
)
```


function to apply the MultMatchCID_spectra() function across compound_ids:

```{r}
run_multmatch <- function(ids, spectra_db, polarity_query, polarity_target,
                          spectrum_type, ms_query, ms_target, minIons = 3,
                          requirePrecursor = TRUE, threshold = 0.9, ppm = 5,
                          tolerance = 0, fragments_resolution = "unit.resolution" ) {

  results <- list()

  for (i in seq_along(ids)) {

    id <- ids[i]
    message("Processing compound_id: ", id)

    res <- tryCatch({

      df <- MultMatchCID_Spectra(
        spectra_db = spectra_db,
        compound_id = id,
        polarity_query = polarity_query,
        polarity_target = polarity_target,
        spectrum_type = spectrum_type,
        ms_level_query = ms_query,
        ms_level_target = ms_target,
        requirePrecursor = requirePrecursor,
        threshold = threshold,
        minIons = minIons,
        fragments_resolution = fragments_resolution,
        ppm = ppm,
        tolerance = tolerance
        
      )

      # if df is a character string (no matches), skip this compound
      if (is.character(df)) {
        message("No matches found for ", id)
        next 
      }

      df$query_compound_id <- id
      df

    }, error = function(e) {
      message("Skipped ", id, " : ", e$message)
      NULL
    })

    if (!is.null(res))
      results[[length(results) + 1]] <- res
  }

  if (length(results) == 0)
    return(NULL)

  do.call(rbind, results)
}
```

Next we call the run_multmatch() function on different MS levels of the 
compounds and the data:

```{r}
res_MS2_MS2 <- run_multmatch(
  compound_ids, spectra_db = dy_sps,
  polarity_query = 0,
  polarity_target = 0,
  spectrum_type = "not_assembled",
  ms_query = 2,
  ms_target = 2,
  threshold = 0.9
)
res_MS2_MS2
```

```{r}
res_MS2_MS3 <- run_multmatch(
  compound_ids, spectra_db = dy_sps,
  polarity_query = 0,
  polarity_target = 0,
  spectrum_type = "not_assembled",
  ms_query = 2,
  ms_target = 3,
  tolerance = 0.6
)
res_MS2_MS3
```


```{r}
res_MS3_MS2 <- run_multmatch(
  compound_ids, spectra_db = dy_sps,
  polarity_query = 0,
  polarity_target = 0,
  spectrum_type = "not_assembled",
  ms_query = 3,
  ms_target = 2,
  tolerance = 0.6
)
res_MS3_MS2
```

```{r}
res_MS3_MS3 <- run_multmatch(
  compound_ids, spectra_db = dy_sps,
  polarity_query = 0,
  polarity_target = 0,
  spectrum_type = "not_assembled",
  ms_query = 3,
  ms_target = 3,
  tolerance = 0.6
)
res_MS3_MS3
```


```{r}
res_MS4_MS2 <- run_multmatch(
  compound_ids, spectra_db = dy_sps,
  polarity_query = 0,
  polarity_target = 0,
  spectrum_type = "not_assembled",
  ms_query = 4,
  ms_target = 2,
  tolerance = 0.6
)
res_MS4_MS2
```

```{r}
res_MS4_MS3 <- run_multmatch(
  compound_ids, spectra_db = dy_sps,
  polarity_query = 0,
  polarity_target = 0,
  spectrum_type = "not_assembled",
  ms_query = 4,
  ms_target = 3,
  tolerance = 0.6
)
res_MS4_MS3
```

```{r}


# List of your result objects
res_list <- list(
  MS2_vs_MS2 = res_MS2_MS2,
  MS2_vs_MS3 = res_MS2_MS3,
  MS3_vs_MS2 = res_MS3_MS2,
  MS3_vs_MS3 = res_MS3_MS3,
  MS4_vs_MS2 = res_MS4_MS2,
  MS4_vs_MS3 = res_MS4_MS3
)

# Concatenate data.frames
all_results <- bind_rows(
  lapply(names(res_list), function(nm) {
    df <- res_list[[nm]]
    if (!is.null(df)) {
      df$comparison <- nm   
      return(df)
    }
    NULL
  })
)

# Save to one CSV
# Create a new workbook
wb <- createWorkbook()

# Add a sheet for the combined results
addWorksheet(wb, "All_Results")

# Write the data.frame to the sheet
writeData(wb, "All_Results", all_results)

# Save the workbook
saveWorkbook(wb, here("data", "FTMSneg_multmatch.xlsx"), overwrite = TRUE)
```

