---
title: "compound_id of identified compounds"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

In this Quarto document we use the MultMatchCID_Spectra() function to extract 
the matches of the flax qtof compounds, that were aligned with the ftms flax 
data, with the dynlib qtof database.


```{r}
#| message: false
#' Load required libraries
library(here)
library(Spectra)
library(CompoundDb)
library(MetaboAnnotation)
library(DBI)
library(RSQLite)
library(devtools)
library(openxlsx)
library(here)
library(writexl)

#` Uploading RDynLib: Load all functions of the RDynlib directory
load_all()
```

## convert the "DYL_FLAX_QTOFneg.sqlite" database to a spectra object

Convert the "DYL_FLAX_QTOFneg.sqlite" database to a spectra object to use it in 
MultmatchCID_spectra() function:

```{r}

dbf <- here("data", "DYL_FLAX_QTOFneg.sqlite")
db <- dbConnect(SQLite(), dbf)
dbListTables(db)
#dbExecute(db, "DROP TABLE IF EXISTS feature_matrix ")
experiment_df <- dbGetQuery(db, "SELECT * FROM experiment")

#here we remove "x", "x.1" columns from experiment table 
experiment_df <- experiment_df[, !names(experiment_df) %in% c("x", "x.1")]
dbExecute(db, "DROP TABLE experiment")
dbWriteTable(db, "experiment", experiment_df)

#the columns have been successfully deleted
dbGetQuery(db, "PRAGMA table_info(experiment)")
dbDisconnect(db)
#we convert the sql database to a Compdb object
cdb <- ExpCompDb(dbf)
#we convert the Compdb object to a spectra object
dy_sps <- Spectra(cdb)
```

```{r}
df <- spectraData(dy_sps, columns = c("msLevel", "compound_id", "expid", "retention_time"))
df[df$compound_id == "45872", ]
```

test
```{r}


cmp <- MultMatchCID_Spectra(
  dy_sps,
  7704,
  ms_level_query = 2,
  ms_level_target = 2,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0,
  ppm = 500,
  minIons = 5,
  fragments_resolution = "high.resolution",
  requirePrecursor = F,
)
cmp
```

qtof compounds that were aligned with the ftms flax data
```{r}
compound_ids <- c(
  45872, 45849, 45918, 45820, 45922, 45885, 45866, 45847,
  45850, 45909, 45910, 45848, 45902, 45923, 45809,
  45905, 45879, 45906, 45878, 45858
)
```

function to apply the MultMatchCID_spectra() function across compound_ids:
```{r}
run_multmatch <- function(ids, spectra_db, polarity_query, polarity_target,
                          spectrum_type_query , ms_query, ms_target, minIons = 5,
                          requirePrecursor = TRUE, threshold = 0.8, ppm = 10,
                          tolerance = 0, fragments_resolution = "high.resolution",
                          spectrum_type_target = "NA") {

  results <- list()

  for (i in seq_along(ids)) {

    id <- ids[i]
    message("Processing compound_id: ", id)

    res <- tryCatch({

      df <- MultMatchCID_Spectra(
        spectra_db = spectra_db,
        compound_id = id,
        polarity_query = polarity_query,
        polarity_target = polarity_target,
        spectrum_type_query  = spectrum_type_query,
        spectrum_type_target = spectrum_type_target,
        ms_level_query = ms_query,
        ms_level_target = ms_target,
        requirePrecursor = requirePrecursor,
        threshold = threshold,
        minIons = minIons,
        fragments_resolution = fragments_resolution,
        ppm = ppm,
        tolerance = tolerance
        
      )

      # if df is a character string (no matches), skip this compound
      if (is.character(df)) {
        message("No matches found for ", id)
        next 
      }

      df$query_compound_id <- id
      df

    }, error = function(e) {
      message("Skipped ", id, " : ", e$message)
      NULL
    })

    if (!is.null(res))
      results[[length(results) + 1]] <- res
  }

  if (length(results) == 0)
    return(NULL)

  do.call(rbind, results)
}
```

print the available spectrum_type values in the database:
```{r}
unique(spectraData(dy_sps, "spectrum_type")[,1])
```

run the function using the merged energies as a spectrum type:
```{r}
res_merged <- run_multmatch(
  compound_ids, spectra_db = dy_sps,
  spectrum_type_query  = "merged_MSMS_all_energies",
  fragments_resolution = "unit.resolution",
  polarity_query = 0,
  polarity_target = 0,
  ms_query = 2,
  ms_target = 2,
  threshold = 0.8,
  minIons = 3,
  ppm = 10,
  tolerance = 0,
  requirePrecursor = T
)
res_merged
```

run the function using the single energy as a spectrum type:
```{r}
res_single <- run_multmatch(
  compound_ids, spectra_db = dy_sps,
  spectrum_type_query = "intersect_single_energy",
  fragments_resolution = "unit.resolution",
  polarity_query = 0,
  polarity_target = 0,
  ms_query = 2,
  ms_target = 2,
  threshold = 0.8,
  ppm = 10,
  tolerance = 0,
  requirePrecursor = T
)
res_single
```

```{r}
unique(res_single$target_compound_id)
```

```{r}
unique(res_single$query_compound_id)
```


```{r}


res_merged$spectrum_type <- "merged_MSMS_all_energies"
res_single$spectrum_type <- "intersect_single_energy"

res_all <- rbind(res_merged, res_single)


write_xlsx(res_all, here("data", "qtofneg_multmatch.xlsx"))
```



```{r}
cid <- spectraData(dy_sps, columns = "compound_id")$compound_id
idx <- which(cid == "4359")
nrow(peaksData(dy_sps[idx])[[1]])
idx
peaksData(dy_sps[idx])[[1]]
```


```{r}
cid <- spectraData(dy_sps, columns = "compound_id")$compound_id
idx <- which(cid == "7704")
nrow(peaksData(dy_sps[idx])[[1]])
idx
peaksData(dy_sps[idx])[[1]]
```
```{r}

```

