---
title: "workflow for downloading and querying the maize DynLib sql data"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe"
---

# Overview

This Quarto document demonstarate how to download the maize DynLib databases  
from zenodo, and how to query the database using SQL commands.
The downloaded databases are acquiered from ftms negative and positive, 
and qtof negative and positive.

## Imprort libraries

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| 
library(devtools)
library(DBI)
library(RSQLite)
library(here)
# TODO: Replace with library(RDynLib) once the package is finalized.
load_all()
```

#this part just before putting the data on zenodo (we have to remove it afterwards)
```{r}
# maize FTneg
maize_ftneg <- here("data", "maize_ftneg.sqlite")

con_mftneg <- dbConnect(RSQLite::SQLite(), maize_ftneg)

#maize FTpos
maize_ftpos <- here("data", "maize_ftpos.sqlite")

con_mftpos <- dbConnect(RSQLite::SQLite(), maize_ftpos)

# maize QTOFneg
maize_qtofneg <- here("data", "maize_qtofneg.sqlite")

con_mqt <- dbConnect(RSQLite::SQLite(), maize_qtofneg)

#maize QTOFpos
#TODO :add the experiment numbers
maize_qtofpos <- here("data", "maize_qtofpos.sqlite")

con_mqtpos <- dbConnect(RSQLite::SQLite(), maize_qtofpos)

```


## Download the databases from zenodo
```{r}
# maize_FTneg 
#RDynlib function to download the database from zenodo
get_dynlib_zenodo(
  doi = ,
  file = "maize_ftmsneg.sqlite",
  destdir = "data"
)
#path of the output database
maize_ftneg <- here("data", "maize_ftneg.sqlite")
#SQL connection to the database
con_mftneg <- dbConnect(RSQLite::SQLite(), maize_ftneg)

#maize FTpos
#RDynlib function to download the database from zenodo
get_dynlib_zenodo(
  doi = ,
  file = "maize_ftmspos.sqlite",
  destdir = "data"
)
#path of the output database
maize_ftpos <- here("data", "maize_ftpos.sqlite")
#SQL connection to the database
con_mftpos <- dbConnect(RSQLite::SQLite(), maize_ftpos)


# maize_QTOFneg
get_dynlib_zenodo(
  doi = ,
  file = "maize_qtofneg.sqlite",
  destdir = "data"
)
#path of the output database
maize_qtofneg <- here("data", "maize_qtofneg.sqlite")
#SQL connection to the database
con_mqt <- dbConnect(RSQLite::SQLite(), maize_qtofneg)

#maize QTOFpos
get_dynlib_zenodo(
  doi = ,
  file = "maize_qtofpos.sqlite",
  destdir = "data"
)
#path of the output database
maize_qtofpos <- here("data", "maize_qtofpos.sqlite")
#SQL connection to the database
con_mqtpos <- dbConnect(RSQLite::SQLite(), maize_qtofpos)

```

## Querying the databases

#Number of experiments in each database:

```{r}
#maize ft neg
dbGetQuery(con_mftneg, "select count(*) from experiment")

```


```{r}
#maize ft pos
dbGetQuery(con_mftpos, "select count(*) from experiment")
```


```{r}
#maize qt neg
dbGetQuery(con_mqt, "select count(*) from experiment")
```

```{r}
#maize qt pos
dbGetQuery(con_mqtpos, "select count(*) from experiment")
```


#compounds number in each database
In maize data each feature has a unique compound_id, so the number of compounds
reported below corresponds to the number of features.

```{r}
#maize ft neg
dbGetQuery(con_mftneg, "select count(*) from ms_compound")

```
```{r}
#maize ft pos
dbGetQuery(con_mftpos, "select count(*) from ms_compound")
```
```{r}
#maize qt neg
dbGetQuery(con_mqt, "select count(*) from ms_compound")
```

```{r}
#maize qt pos
dbGetQuery(con_mqtpos, "select count(*) from ms_compound")
```
#Counting the spectra numbers in each database

```{r}
#maize ft neg
dbGetQuery(con_mftneg, "select count(*) from msms_spectrum")

```


```{r}
#maize ft pos
dbGetQuery(con_mftpos, "select count(*) from msms_spectrum")
```


```{r}
#maize qt neg
dbGetQuery(con_mqt, "select count(*) from msms_spectrum")
```

```{r}
#maize qt pos
dbGetQuery(con_mqtpos, "select count(*) from msms_spectrum")
```

#The corresponding rows of a given experiment

```{r}
dbGetQuery(con_mqtpos, "select * from ms_compound where expid = 3")
```

```{r}
dbGetQuery(con_mftneg, "select * from experiment where expid IN (199, 230);")
```
#The corresponding row of a given compound

```{r}
dbGetQuery(con_mftneg, "select * from ms_compound where compound_id = 63668")
```
#The corresponding spectrum(s) of a given compound

```{r}
dbGetQuery(con_mftneg, "select * from msms_spectrum where compound_id = 63668")
```

#the corresponding peaks of  a given spectum

```{r}
dbGetQuery(con_mftneg, "select * from msms_spectrum_peak where spectrum_id = 157135")
```

