---
title: "workflow to test the MultMatchCID_Spectra() function on ftms data"
format: html
  server: shiny
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This quarto document represents the workflow for calling the`MultMatchCID_Spectra()`
from the [MultMatchCID_Spectra.R](https://github.com/rebdau/RDynLib_SQL/blob/Ahlam/R/MultMatchCID_Spectra.R)
file to calculate the similarity between  a given compound_id from the DynLib 
FTMS negative database and the other compounds from the same database.

# Required packages

This project follows an R package layout as follows:

The main project folder is:

RDynLib_SQL/

Within this directory, the structure is organized as follows:

RDynLib_SQL/ ├── data/ ├── R/ └── vignettes/

### `data/`

Contains the SQLite database files used in the workflow 
(e.g., `ftms_flax_neg.sqlite`).

### `R/`

Contains the R source files (`.R`) that define the package functions and core
implementation.

### `vignettes/`

Contains Quarto (`.qmd`) files that document the workflow and provide 
reproducible examples.

In this workflow, the working directory is set to the project root:

RDynLib_SQL/

## Loading required packages

First, install the new CompoundDb version:

```{r}
BiocManager::install("RforMassSpectrometry/CompoundDb", ref = "jomain", force = TRUE)
```

```{r}
library(here)
library(Spectra)
library(CompoundDb)
library(MetaboAnnotation)
library(DBI)
library(RSQLite)
library(devtools)

#` Uploading RDynLibSQL: Load all functions of the RDynlib directory
load_all()
```

## convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

Next we convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

```{r}

dbf <- here("data", "DYL_FLAX_FTMSneg.sqlite")
db <- dbConnect(SQLite(), dbf)
dbListTables(db)
#dbExecute(db, "DROP TABLE IF EXISTS feature_matrix ")
experiment_df <- dbGetQuery(db, "SELECT * FROM experiment")

#here we remove "x", "x.1" columns from experiment table 
experiment_df <- experiment_df[, !names(experiment_df) %in% c("x", "x.1")]
dbExecute(db, "DROP TABLE experiment")
dbWriteTable(db, "experiment", experiment_df)

#the columns have been successfully deleted
dbGetQuery(db, "PRAGMA table_info(experiment)")
dbDisconnect(db)
#we convert the sql database to a Compdb object
cdb <- ExpCompDb(dbf)
#we convert the Compdb object to a spectra object
dy_sps <- Spectra(cdb)
```

```{r}
spectraData(dy_sps, columns = c("retention_time"))
```



#test the 'MultMatchCID_Spectra()' function to return the matches of a given 
#compound_id with the whole database(where the user should enter the spectrum 
#type and mslevel )

```{r}
cmp_matches <- MultMatchCID_Spectra(
  dy_sps,
  100988,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0,
  ppm = 10,
  minIons = 2,
  fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)


```


```{r}
cmp <- MultMatchCID_Spectra(
  dy_sps,
  101280,
   spectrum_type_query = "not_assembled",
   ms_level_query = 2,
   ms_level_target = 2,
    polarity_query = 0,
  polarity_target = 0,
  tolerance = 0,
  ppm = 10,
  minIons = ,
    fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
cmp
```



print the results:

```{r}
print(cmp_matches)
nrow(cmp_matches)
```


#test
```{r}
sum(round(test2$`m/z`,2) != 355.07)
```



save the resulting object:

```{r}
save(cmp_matches, file = "data/cmp_matches.RData")
```

```{r}
load("data/cmp_matches.RData")
```

