---
title: "workflow to test the MultMatchCID_Spectra() function on ftms data"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This quarto document represents the workflow for calling the`MultMatchCID_Spectra()`
from the [MultMatchCID_Spectra.R](https://github.com/rebdau/RDynLib_SQL/blob/Ahlam/R/MultMatchCID_Spectra.R)
file to calculate the similarity between  a given compound_id from the DynLib 
FTMS negative database and the other compounds from the same database.
Also, we show how to create a molecular network from a given compound_id, and 
finaly we export the sql databases to a mgf files for further analyses.

# Required packages

This project follows an R package layout as follows:

The main project folder is:

RDynLib_SQL/

Within this directory, the structure is organized as follows:

RDynLib_SQL/ ├── data/ ├── R/ └── vignettes/

### `data/`

Contains the SQLite database files used in the workflow 
(e.g., `ftms_flax_neg.sqlite`).

### `R/`

Contains the R source files (`.R`) that define the package functions and core
implementation.

### `vignettes/`

Contains Quarto (`.qmd`) files that document the workflow and provide 
reproducible examples.

In this workflow, the working directory is set to the project root:

RDynLib_SQL/

## Loading required packages

First, install the new CompoundDb version:

```{r}
BiocManager::install("RforMassSpectrometry/CompoundDb", ref = "jomain", force = TRUE)
```

```{r}
library(here)
library(Spectra)
library(CompoundDb)
library(MetaboAnnotation)
library(DBI)
library(RSQLite)
library(devtools)
library(here)
library(dplyr)
library(stringr)
#` Uploading RDynLibSQL: Load all functions of the RDynlib directory
load_all()
```

## convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

Next we convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

```{r}
dbf <- here("data", "DYL_flax_herba_FTMSneg.sqlite")
db <- dbConnect(SQLite(), dbf)
dy_sps <- sql_to_spectra(dbf)
```

```{r}
spectraData(dy_sps, columns = c("retention_time"))
```

#test the 'MultMatchCID_Spectra()' function to return the matches of a given 
#compound_id with the whole database(where the user should enter the spectrum 
#type and mslevel )

We start by matching the 101280 compound with the database:

```{r}
mt_3_3 <- MultMatchCID_Spectra(
  dy_sps,
  101280,
  spectrum_type_query = "not_assembled",
  spectrum_type_target = "not_assembled",
  ms_level_query = 3,
  ms_level_target = 3,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0.6,
  ppm = 100,
  minIons = 2,
  threshold = 0.8,
  fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
mt_3_3
```
The MS3 vs MS3 returns 71 match in the database.


```{r}
mt_2_2 <- MultMatchCID_Spectra(
  dy_sps,
  101280,
  spectrum_type_query = "not_assembled",
  spectrum_type_target = "not_assembled",
  ms_level_query = 2,
  ms_level_target = 2,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0.6,
  ppm = 100,
  minIons = 2,
  threshold = 0.8,
  fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
mt_2_2
```

The MS2 vs MS2 returns 28 match in the database.

```{r}
mt_4_3 <- MultMatchCID_Spectra(
  dy_sps,
  101280,
  spectrum_type_query = "not_assembled",
  spectrum_type_target = "not_assembled",
  ms_level_query = 4,
  ms_level_target = 3,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0.6,
  ppm = 100,
  minIons = 2,
  threshold = 0.8,
  fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
mt_4_3
```

The MS4 vs MS3 returns 13 match in the database.

```{r}
mt_4_4 <- MultMatchCID_Spectra(
  dy_sps,
  101280,
  spectrum_type_query = "not_assembled",
  spectrum_type_target = "not_assembled",
  ms_level_query = 4,
  ms_level_target = 4,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0.6,
  ppm = 100,
  minIons = 2,
  threshold = 0.8,
  fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
mt_4_4
```

The MS4 vs MS3 returns 1 match in the database.

look for a given name
```{r}
mt_3_3 %>%
  filter(str_detect(target_name, regex("herbacetin 3,8-di-o-beta-d-glucopyranoside ID AM", ignore_case = TRUE)))
```
we look for the ms_compound table of a given compound_id of HDG-succinate-HDG:
```{r}
cmp <- get_compound(dbf, 101099)
cmp
```

First we fill in the cspp table:

```{r}
ft_path <- here("data", "DYL_flax_herba_FTMSneg.sqlite")
alg <- here("data", "assoc_flax_63.txt")
cspp_flax <- here("data","cspp_flax.txt")
charge_file <- here("data", "ftms_all_compounds_RD.xlsx")
```


```{r}
cspp.tot_SQL(ft_path, 266, mzerr = 0.01,
                         cspp = cspp_ft, peakwidth = 0.1, IntThres  = 0, mz_tol = 0.5,
                         dot_thresh = 0, spectrum_type = "assembled",  
                         min_common = 0, charge_file = charge_file)
```


Here we show the cspp network of that compound
```{r}
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 100988, min = 0,
                nr_of_seq = 3, thr1 = 0.1, thr2 = 0.2, thr3 = 0.3)
```

Next, we print the compounds from qtof neg that have been aligned with
HDG-succinate-HDG:

```{r}
qt_path <- here("data","DYL_FLAX_QTOFneg.sqlite")
```

```{r}
getAlignedCompids_SQL(100988, "DYL_FLAX_FTMSneg.sqlite", alg )
```

fill in the gnps_add table
```{r}
conv.GNPS_SQL(sql_path = qt_path, expid= 52, mzerr = 0.015,
              min = 5, adduct = 60.0211, thr1 = 3, thr2 = 0.1, thr3 = 0.4,
              spectrum_type = "merged_MSMS_all_energies", IntThres = 5)
```

fill in the cspp_add table
```{r}
cspp.tot_SQL(qt_path, 52, mzerr = 0.01,
                         cspp = cspp_flax, peakwidth = 0.2, IntThres  = 0, mz_tol = 0.5,
                         dot_thresh = 0.7, spectrum_type = "intersect_single_energy",  
                         min_common = 3, charge_file = charge_file)
```


Here we print the ms_compound table of that compound from qtof database:
```{r}
cmp_qt <- get_compound(qt_path, 45890)
cmp_qt
```

Finaly, we show the GNPS network of that compound
```{r}
Overall.net_SQL(sql_path = qt_path, exp.id = 52, dbkey = 45890, min = 0,
                nr_of_seq = 3, thr1 = 0.1, thr2 = 0.2, thr3 = 0.3)
```


Next, we call the function to export the sql database to an mgf format that 
could be used for further analysis and comparisons with public databases 
available on public platforms like: GNPS...
We start by converting ftms data
```{r}
#| message: false
#| results: "hide"
#| warning: false
ft_assembled <- here("data", "ft_assembled.mgf")
sql_to_mgf(ft_path, expid = 266, spectrum_type = "assembled", 
           output_file = ft_assembled) 
```

```{r}
#| message: false
#| results: "hide"
#| warning: false
ft_assembled <- here("data", "ft_MS2.mgf")
sql_to_mgf(ft_path, expid = 266, spectrum_type = "not_assembled", 
           msLevel = 2, output_file = ft_assembled) 
```

Then, we convert qtof data:
```{r}
#| message: false
#| results: "hide"
#| warning: false
qt_assembled <- here("data", "qt_assembled.mgf")
sql_to_mgf(qt_path, expid = 52, spectrum_type = "intersect_single_energy", 
           output_file = qt_assembled) 
```


```{r}
#| message: false
#| results: "hide"
#| warning: false
qt_assembled <- here("data", "qt_merged.mgf")
sql_to_mgf(qt_path, expid = 52, spectrum_type = "merged_MSMS_all_energies", 
           output_file = qt_assembled)
```

