---
title: "workflow to test the similarity_RDynlib() function on qtof data"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This quarto document represents the workflow for calling the `similarity_RDynlib()` from the [similarity_functions.R](https://github.com/rebdau/RDynLib_SQL/blob/Ahlam/R/similarity_functions.R) file to calculate the similarity between QTOF flax data and the RDynLib QTOFdatabase.

# Required packages

```{r}
library(here)
library(Spectra)
library(CompoundDb)
library(MetaboAnnotation)
library(DBI)
library(RSQLite)
library(devtools)

#` Uploading RDynLibSQL: Load all functions of the RDynlib directory
load_all()
```

Then, install the new CompoundDb version:

```{r}
BiocManager::install("RforMassSpectrometry/CompoundDb", ref = "jomain", force = TRUE)
```

```{r}
ExpCompDb <- function(x) {
    cdb <- CompDb(x)
    CompoundDb::addJoinDefinition(cdb, "ms_compound", "experiment", "expid", "expid")
}
```

Next we convert the "DyLib_QTOF_neg2.sqlite" database to a spectra object

```{r}
library(DBI)
library(RSQLite)
dbf <- here("data", "DyLib_QTOF_neg.sqlite")
db <- dbConnect(SQLite(), dbf)
dbExecute(db, "DROP TABLE IF EXISTS feature_matrix ")
experiment_df <- dbGetQuery(db, "SELECT * FROM experiment")

#here we remove "x", "x.1" columns from experiment table 
experiment_df <- experiment_df[, !names(experiment_df) %in% c("x", "x.1")]
dbExecute(db, "DROP TABLE experiment")
dbWriteTable(db, "experiment", experiment_df)

#the columns have been successfully deleted
dbGetQuery(db, "PRAGMA table_info(experiment)")
dbDisconnect(db)
#we convert the sql database to a Compdb object
cdb <- ExpCompDb(dbf)
#we convert the Compdb object to a spectra object
qt_dy_sps <- Spectra(cdb)
```

Here we load the QTOF of flax seed database (it's already a spectra object), you can load it from: [qtof_filtering.qmd](https://github.com/rebdau/RDynLib_qtof/blob/ahlam/qtof_filtering.qmd)

```{r}
load(here("data/database_QTOF_neg","qtof_flax_final.RData"))
qtof_neg <- qtof_flax_final
spectraData(qtof_neg)
```

Next we calculate the similarity score between the dynlib and flax seed QTOF negative integuments.(intersect_single_energy)

```{r}
qtof_matches <- similarity_RDynlib_match(
  qtof_neg,
  qt_dy_sps,
  
  polarity_query = 0,
  polarity_target = 0,
  fragments_resolution = "high.resolution",
  threshold = 0.8,
  ppm = 10,
  tolerance = 0,
  requirePrecursor = FALSE
)
```

print the similarity matrix and the number of rows (matches)

```{r}
print(qtof_matches)
nrow(qtof_matches)
```

save the resulting object:

```{r}
save(qtof_matches, file = here("data/database_QTOF_neg", "qtof_matches.RData"))
```
