---
title: "Selection of representative MS2 spectra per feature"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
editor: 
  markdown: 
    wrap: 72
---

## Introdcution

In this quarto document we filtered the QTOF neg of flax seed data in
order to retain one representative MS/MS spectrum per feature. The steps
and parameters were tested in
`qtof_filtering_exploreknownfragmentation.qmd`

We create the precursorPurity and PrecursorIntensity columns. We
centroid the MS2 spectra. We select the 10 % spectra with highest
precursor purity. We combine spectra to obtain one representative
spectrum per feature, and finally keep for each spectrum the 100 peaks
with highest intensities.

## Data upload

```{r, echo = FALSE}
library(Spectra)
library(xcms)
library(MsExperiment)
library(S4Vectors)
library(BiocParallel)
library(Rdisop)
library(stringr)
library(dplyr)

register(SerialParam())

pth <- getwd()
pth

```

Load the qtof final object after preprocessing from
ftms_preprocessing.qmd file, and convert it to a spectra object

```{r}
load(file.path(pth, "data/qtof.RData"))
```

In case of changing the path to the original mzml files used to create the qtof 
MsExperiment object please run this chunk and replace the path with the actual
path of your mzml file:
```{r}
dataStorageBasePath(qtof@spectra) <- file.path(pth, "data/mzml") 
```

Convert to a spectra object
```{r}
qtof_neg <- spectra(qtof)
```


## Calculate Precursor purity

We next calculate the precursor purity, by calling the
`precursorPurity()` function from the spectra package, for each MS2
spectrum as well as the precursor intensity (since that is not provided
in the original mzML files).

### Function

Note: this function was added to the *Spectra* package with pull request
[#359](https://github.com/rformassspectrometry/Spectra/pull/359) and will be
available in the *Spectra* package starting with Bioconductor release 3.22
(October 2025). In the meantime, the function can be loaded below.

```{r}
precursorPurity <- function(x, tolerance = 0.3, ppm = 10) {
   # if (is.unsorted(rtime(x)))
    #    stop("Spectra with data origin ", dataOrigin(x[1L]),
     #        " are not increasingly sorted by retention time.")
    ms2_idx <- which(msLevel(x) == 2L)
    ms1_idx <- which(msLevel(x) == 1L)
    ratios <- rep(NA_real_, length(x))
    pmzs <- precursorMz(x)
    for (i in ms2_idx) {
        ratio <- NA_real_
        j <- ms1_idx[ms1_idx < i]

        if (length(j)) {
            j <- j[length(j)]
            s1 <- filterMzRange(
                x[j], c(pmzs[i] + c(-1, 1) * (tolerance + pmzs[i] * ppm / 1e6)))

            pks <- peaksData(s1, c("mz", "intensity"))[[1L]]

            if (nrow(pks)) {
                intensities <- pks[, 2L]
                ratio <- max(intensities) / sum(intensities)
            }
        }

        ratios[i] <- ratio
    }

    ratios
}


```

### Calculation

```{r}
qtof_neg$precursorPurity <- precursorPurity(qtof_neg, tolerance = 0.3, ppm = 10)
qtof_neg$precursorIntensity <- estimatePrecursorIntensity(
    qtof_neg, tolerance = 0.3, ppm = 10)
qtof@spectra <- qtof_neg
```

The distribution of the estimated precursor purity and intensity are shown
below.

```{r}
#| fig-width: 10
#| fig-height: 5
par(mfrow = c(2, 1))
plot(density(qtof_neg$precursorPurity, na.rm = TRUE),
     ylab = "Precursor purity", main = "")
grid()
plot(density(qtof_neg$precursorIntensity, na.rm = TRUE),
     ylab = "Precursor intensity", main = "")
grid()
```

### Save XcmsExperiment object with precursorPurity

```{r}
save(qtof, file = file.path(pth, "data", "qtof_precpur.RData"))
```

### Load XcmsExperiment object with precursorPurity

```{r}
load(file = "data/qtof_precpur.RData")
```

## Associate MS2 spectra to features

We next extract all MS2 spectra that can be associated to the MS1
feature:

```{r}
feature_spectra <- featureSpectra(dropFilledChromPeaks(qtof), msLevel = 2,
                                  skipFilled = TRUE)
```

For only `r length(unique(feature_spectra$feature_id))` of in total
`r nrow(featureDefinitions(qtof))` features MS2 spectra were defined. A summary
of the number of MS2 spectra per feature is shown below:

```{r}
table(feature_spectra$feature_id) |>
    quantile()
```

A relatively large number of MS2 spectra was assigned to each feature. A
summary of the number of fragment peaks per spectrum is given below.

```{r}
lengths(feature_spectra) |>
    quantile()
```

The number of peaks per spectrum is unexpectedly large.

## Centroiding the MS2 data

To reduce the number of fragment peaks we perform a second centroiding of the
spectra.

```{r}
feature_spectra_centro <- Spectra::pickPeaks(feature_spectra,
                                             halfWindowSize = 2L,
                                             k = 2L, threshold = 0.5)
```

The fragment peak counts of the spectra after centroiding is:

```{r}
lengths(feature_spectra_centro) |>
    quantile()
```

The number of peaks was reduced - is however still large.

statistics of the features with at least one assigned MS2 spectrum and Total 
number of detected features
```{r}
# Percentage of features with at least one MS2 spectrum
percentage_assigned <- 
    100 * length(unique(feature_spectra$feature_id)) /
    nrow(featureDefinitions(qtof))
percentage_assigned
```


## Verify if MS2 Spectra are linked to more than one feature

```{r}
scanIndex.per.ori <- split(feature_spectra_centro$scanIndex,
                           feature_spectra_centro$dataOrigin)
```

```{r}
ms2_id <- paste0(feature_spectra_centro$dataOrigin, "_", feature_spectra_centro$acquisitionNum)
ms2_table <- table(ms2_id)
ms2_mult <- names(ms2_table)[ms2_table > 1]
length(ms2_mult)
```

There is `r length(ms2_mult)` duplicated assignment.

In case of the length of `ms2_mult` not null, this file
[Resolve_multiple_assignment.R](https://github.com/rebdau/RDynLib_qtof/blob/ahlam/Resolve_multiple_assignment.R)
contains an algorithm that remove the duplicated assignments


#case 1: If your data were acquired using only a single collision energy

## Precursor purity

### Removing a fraction of the spectra based on precursor purity

We next remove fragment spectra with a low precursor purity. For this we group
fragment spectra by feature ID and remove, for each group,
the 10% of spectra with the lowest precursor purity.

```{r}
# extract the spectra data as a dataframe
filenames <- feature_spectra$dataOrigin

purity_df <- as.data.frame(spectraData(feature_spectra_centro)) %>%
  mutate(
    index = row_number(),
    f = feature_id
  )

# 10% lowest
keep_idx <- purity_df %>%
  group_by(f) %>%
  arrange(precursorPurity) %>%
  mutate(
    n_rows = n(),
    cutoff = ceiling(n_rows * 0.10),
    row_number = row_number()
  ) %>%
  filter(row_number > cutoff) %>%
  pull(index)
ms2_qtof_precpur <- feature_spectra_centro[keep_idx]

#number of spectra:
dim(spectraData(ms2_qtof_precpur))[1]

#number of peaks per spectrum:
counts <- sapply(peaksData(ms2_qtof_precpur), function(x) {
  dim(x)[1]
})
summary(counts)
hist(counts)
```

Comparing the precursor purity before and after filtering.

```{r}
quantile(feature_spectra_centro$precursorPurity, na.rm = TRUE)
quantile(ms2_qtof_precpur$precursorPurity, na.rm = TRUE)
```

The global precursor purities slightly improved. After filtering,
`r length(ms2_qtof_precpur)` MS2 spectra from the in total
`r length(feature_spectra_centro)` are retained.

## Merge spectra within feature_id
Create a single spectrum for each feature by combining all spectra associated 
with the same feature_id , keeping only fragments that are present in minimum 
`minprop` of spectra. We define that peaks need to be present in at least 50% 
of the spectra, otherwise they are considered noise.

```{r}
# grouping by feature_id 
f <- ms2_qtof_precpur$feature_id

ms2_qtof_ftdda <- combineSpectra(
    ms2_qtof_precpur, ppm = 5, peaks = "intersect", weighted = TRUE,
    minProp = 0.5, intensityFun = mean, f = f,
    p = rep(1L, length(ms2_qtof_precpur)))
```

Combining reduced the number of MS2 spectra from `r length(ms2_qtof_precpur)`
to `r length(ms2_qtof_ftdda)`.

The number of peaks of the combined MS2 spectra is summarized below

```{r}
lengths(ms2_qtof_ftdda) |>
    quantile()
```

There are thus `sum(lengths(ms2_qtof_ftdda) <= 1)` spectra with one or no peak.
We need to remove the spectra without available peaks.

```{r}
ms2_qtof_ftdda <- ms2_qtof_ftdda[lengths(ms2_qtof_ftdda) > 0]
```


#case 2: If the data were acquired using different collision energies

## Precursor purity

### Removing a fraction of the spectra based on precursor purity

We next remove fragment spectra with a low precursor purity. For this we group
fragment spectra by feature ID and collision energy and remove, for each group,
the 10% of spectra with the lowest precursor purity.

```{r}
# extract the spectra data as a dataframe
filenames <- feature_spectra$dataOrigin
dda <- str_extract(filenames, "DDA[0-9]")

purity_df <- as.data.frame(spectraData(feature_spectra_centro)) %>%
  mutate(
    index = row_number(),
    f = paste(feature_id, "|", dda)
  )

# 10% lowest
keep_idx <- purity_df %>%
  group_by(f) %>%
  arrange(precursorPurity) %>%
  mutate(
    n_rows = n(),
    cutoff = ceiling(n_rows * 0.10),
    row_number = row_number()
  ) %>%
  filter(row_number > cutoff) %>%
  pull(index)
ms2_qtof_precpur <- feature_spectra_centro[keep_idx]

#number of spectra:
dim(spectraData(ms2_qtof_precpur))[1]

#number of peaks per spectrum:
counts <- sapply(peaksData(ms2_qtof_precpur), function(x) {
  dim(x)[1]
})
summary(counts)
hist(counts)
```

Comparing the precursor purity before and after filtering.

```{r}
quantile(feature_spectra_centro$precursorPurity, na.rm = TRUE)
quantile(ms2_qtof_precpur$precursorPurity, na.rm = TRUE)
```

The global precursor purities slightly improved. After filtering,
`r length(ms2_qtof_precpur)` MS2 spectra from the in total
`r length(feature_spectra_centro)` are retained.

## Merge spectra within feature_id - collision energy groups
create a single spectrum for each feature per collision energy, keeping
only fragments that are present in minimum `minprop` of spectra. We
define that peaks need to be present in at least 50% of the spectra,
otherwise they are considered noise.

```{r}
# grouping by feature_id and collision energy
filenames <- ms2_qtof_precpur$dataOrigin

# Extract the DDA group (e.g. DDA1, DDA2, ...)
dda <- str_extract(filenames, "DDA[0-9]")

# Create the grouping factor:
# spectra are combined only when they have the same feature_id
# AND belong to the same DDA/collision-energy group
f <- paste(ms2_qtof_precpur$feature_id, "|", dda)

ms2_qtof_ftdda <- combineSpectra(
    ms2_qtof_precpur, ppm = 5, peaks = "intersect", weighted = TRUE,
    minProp = 0.5, intensityFun = mean, f = f,
    p = rep(1L, length(ms2_qtof_precpur)))
```

Combining reduced the number of MS2 spectra from `r length(ms2_qtof_precpur)`
to `r length(ms2_qtof_ftdda)`.

# Percentage of features retained after precursor purity filtering
```{r}
length(unique(ms2_qtof_precpur$feature_id)) /
    nrow(featureDefinitions(qtof)) * 100
```


The number of peaks of the combined MS2 spectra is summarized below

```{r}
lengths(ms2_qtof_ftdda) |>
    quantile()
```

There are thus `sum(lengths(ms2_qtof_ftdda) <= 1)` spectra with one or no peak.
We need to remove the spectra without available peaks.

```{r}
ms2_qtof_ftdda <- ms2_qtof_ftdda[lengths(ms2_qtof_ftdda) > 0]
```

# Percentage of features retained after combining spectra by collision energy
```{r}
percentage_after_combining <- 
    100 * length(unique(ms2_qtof_ftdda$feature_id)) /
    nrow(featureDefinitions(qtof))
percentage_after_combining
```

## For both cases: Keep 100 peaks with highest intensity

In contrast to very few peaks, there are also MS2 spectra with a high number of
peaks. For these, we restrict the number of fragment peaks to 100.

```{r}
# Keep only the 100 peaks with the highest intensity
top100 <- function(x, ...) {
    # If fewer than 100 peaks, keep them all
    if (length(x) <= 100) {
        rep(TRUE, length(x))
    } else {
        # Keep TRUE for the 100 highest intensities
        rank(-x, ties.method = "first") <= 100
    }
}

# Apply the filter
ms2_qtof_sep <- filterIntensity(ms2_qtof_ftdda, intensity = top100)

spectraData(ms2_qtof_sep)$spectrum_type <- "same_energy"


```

The number of peaks before and after this filtering are shown below

```{r}
lengths(ms2_qtof_ftdda) |> quantile()
lengths(ms2_qtof_sep) |> quantile()
```

Number of spectra with just one peak

```{r}
sum(lengths(ms2_qtof_sep) == 1)
```


```{r}
save(ms2_qtof_sep, file = file.path(pth, "data", "ms2_qtof_sep.RData"))

```

N.B.: If your data were acquired using only a single collision energy, 
then your final object is ms2_qtof_sep.RData (loaded above).
In this case, do not continue with the rest of the workflow.
If your data were acquired using multiple collision energies, 
load the ms2_qtof_sep.RData object and continue with the following
steps of the workflow.

## Grouping by feature_id for different collision energies

```{r}
# grouping by feature_id for different collision energies
f = ms2_qtof_ftdda$feature_id
ms2_qtof <- combineSpectra(ms2_qtof_ftdda, ppm = 5, peaks = "union",
                           intensityFun = sum,weighted = TRUE, f = f,
                           p = rep(1L, length(ms2_qtof_ftdda)))


```

```{r}
length(unique(unlist(ms2_qtof$feature_id)))

length(unique(spectraData(ms2_qtof)$feature_id))
```

As a result we have `r length(unique(spectraData(ms2_qtof)$feature_id))`
unique feature

### Check number of peaks per spectrum

```{r}

#number of peaks per spectrum
counts <- sapply(peaksData(ms2_qtof), function(x) {
  dim(x)[1]
})

summary(counts)

hist(counts, breaks = 100)
hist(counts, breaks = 100, xlim = c(0,400))
```

## Keep 100 peaks with highest intensity

```{r}
# Keep only the 100 peaks with the highest intensity
top100 <- function(x, ...) {
    # If fewer than 100 peaks, keep them all
    if (length(x) <= 100) {
        rep(TRUE, length(x))
    } else {
        # Keep TRUE for the 100 highest intensities
        rank(-x, ties.method = "first") <= 100
    }
}

# Apply the filter
ms2_qtof_comb <- filterIntensity(ms2_qtof, intensity = top100)
spectraData(ms2_qtof_comb)$spectrum_type <- "all_energies"
```

Here we print the summary and the quantile of the peaks number

```{r}
peak_counts <- sapply(peaksData(ms2_qtof_comb), nrow)
summary(peak_counts)
quantile(peak_counts)
```

save the object

```{r}
save(ms2_qtof_comb, file = file.path(pth, "data", "ms2_qtof_Comb.RData"))
```

Here we save the global object that contains the qtof assembled and
separated, distinguished by spectrum_type column

```{r}
qtof_flax_final <- c(ms2_qtof_sep, ms2_qtof_comb)
save(qtof_flax_final, file = file.path(pth, "data", "qtof_flax_final.RData"))
```

Statistics part:

```{r}
sd <- spectraData(qtof_flax_final) |> as.data.frame()

unique(sd$spectrum_type)

# extract energy level (DDA1/DDA2/DDA3) from the file path
sd$energy <- str_extract(sd$dataOrigin, "DDA[0-9]")

## MS2 spectra count PER ENERGY LEVEL, for spectrum_type == "same_energy"
counts_single_energy <- sd %>%
  filter(msLevel == 2, spectrum_type == "same_energy") %>%
  count(energy, name = "n_ms2_spectra")

counts_single_energy
```


```{r}
## Total MS2 spectra where spectrum_type == "all_energies"
n_merged_ms2 <- sd %>%
  filter(msLevel == 2, spectrum_type == "all_energies") %>%
  nrow()

n_merged_ms2
```


```{r}
counts_merged_total <- sd %>%
  filter(msLevel == 2, spectrum_type == "all_energies") %>%
  summarise(n_ms2_spectra = n())

counts_merged_total


counts_merged_by_energy <- sd %>%
  filter(msLevel == 2, spectrum_type == "all_energies") %>%
  count(energy, name = "n_ms2_spectra")

counts_merged_by_energy
```



## Session information

```{r}
sessionInfo()
```
