---
title: "RDynLib workflow for metabolite structural annotation"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe"
---

# Overview

This workflow illustrates how the **RDynLib infrastructure supports metabolite structural annotation**, from spectral data exploration and comparison to the recording of compound annotations and their provenance in DynLib databases.

A flax seed DynLib database, generated from a new LC-MS experiment containing compounds to be structurally annotated, serves as the **query database** throughout the workflow. It is compared with a maize DynLib database containing previously characterized metabolites and their associated spectra, which serves as the **reference database**.

The workflow progresses from general **spectral similarity searching** against a reference DynLib database (the maize database in this example) and the **export of spectral data for further analysis with external databases and tools**, to detailed comparison and **visualization of individual spectra**, including **matching across different MSⁿ levels**. A key functionality of RDynLib is the **joint exploitation of complementary fragmentation information acquired on different mass spectrometry platforms**, such as ion trap and QTOF instruments, to support structural characterization. Biochemical and spectral relationships are explored through the construction and visualization of **CSPP and GNPS-like molecular networks**.

Finally, the resulting **compound annotation is written to the flaxseed DynLib database**. Along with the structural annotation itself, the evidence and reasoning underlying the characterization are recorded as an **annotation event**, preserving the provenance of the annotation and documenting the justification for the **assigned confidence level**.

# Workflow setup

## Loading required packages

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| 
library(devtools)
library(DBI)
library(RSQLite)
library(here)
library(CompoundDb)
library(MsBackendMgf)
library(scales)
# TODO: Replace with library(RDynLib) once the package is finalized.
load_all()
```

## Retrieving DynLib datasets from Zenodo

The DynLib SQLite databases used in this workflow are provided through Zenodo. Each database contains the features detected in one or more LC-MS experiments together with their associated mass spectral data and experimental metadata and compound annotations. Separate databases are used for each mass spectrometry platform and ionization mode:

-   **FTneg / FTpos**: LTQ-Orbitrap FT-MS data acquired in negative / positive ionization mode, including MSⁿ spectra
-   **QTOFneg / QTOFpos**: QTOF-MS data acquired in negative / positive ionization mode, including MS/MS spectra

```{r}
# TODO: this chunk is for the working document and needs to be replaced by the Zenodo chunk when we will have deposited the databases.
#| message: false
#| results: "hide"
#| warning: false

#big ftmsneg
ftneg_path <- here("data", "DYL_FLAX_herba_FTMSneg.sqlite")
con_ftneg <- dbConnect(RSQLite::SQLite(), ftneg_path)

#big ftmspos
ftpos_path <- here("data", "DyLib_FTMS_pos.sqlite")
con_ftpos <- dbConnect(RSQLite::SQLite(), ftpos_path)

# big QTOFneg
qtofneg_path <- here("data", "DYL_FLAX_QTOFneg.sqlite")
con_qtofneg <- dbConnect(RSQLite::SQLite(), qtofneg_path)

#big qtof pos

qtofpos_path <- here("data", "DyLib_QTOF_pos.sqlite")

# maize FTneg
maize_ftneg <- here("data", "maize_ftneg.sqlite")
filterDynLibDB(
  expid = 191:231,
  input_db = ftneg_path,
  output_db = maize_ftneg
)
con_mftneg <- dbConnect(RSQLite::SQLite(), maize_ftneg)

#maize FTpos
maize_ftpos <- here("data", "maize_ftpos.sqlite")
filterDynLibDB(
  expid = 55:94,
  input_db = ftpos_path,
  output_db = maize_ftpos
)
con_mftpos <- dbConnect(RSQLite::SQLite(), maize_ftpos)

# maize QTOFneg
maize_qtofneg <- here("data", "maize_qtofneg.sqlite")
filterDynLibDB(
  expid = 38,
  input_db = qtofneg_path,
  output_db = maize_qtofneg
)
con_mqt <- dbConnect(RSQLite::SQLite(), maize_qtofneg)

#maize QTOFpos
#TODO :add the experiment numbers
maize_qtofpos <- here("data", "maize_qtofpos.sqlite")
filterDynLibDB(
  expid = 3,
  input_db = qtofpos_path,
  output_db = maize_qtofpos
)
con_mqtpos <- dbConnect(RSQLite::SQLite(), maize_qtofpos)

# flax FTneg
flax_ftneg <- here("data", "flax_ftneg.sqlite")
filterDynLibDB(
  expid = 266,
  input_db = ftneg_path,
  output_db = here("data", "flax_ftneg.sqlite")
)
con_fft <- dbConnect(RSQLite::SQLite(), flax_ftneg)

# flax QTOFneg
flax_qtofneg <- here("data", "flax_qtofneg.sqlite")
filterDynLibDB(
  expid = 52,
  input_db = qtofneg_path,
  output_db = here("data", "flax_qtofneg.sqlite")
)
con_fqt <- dbConnect(RSQLite::SQLite(), flax_qtofneg)

# HDG FTneg 
hdg_ftneg <- here("data", "hdg_ftneg.sqlite")
filterDynLibDB(
  expid = 267,
  input_db = ftneg_path,
  output_db = here("data", "hdg_ftneg.sqlite"))
con_hdgft <- dbConnect(RSQLite::SQLite(), hdg_ftneg)

```

```{r}
# TODO: this chunk downloads from ZENODO and will replace the above chunk when we will have deposited the databases.
#| message: false
#| results: "hide"
#| warning: false
# Import the data from zenodo and create the connections
# maize_FTneg
get_dynlib_zenodo(
  doi = ,
  file = "maize_ftmsneg.sqlite",
  destdir = "data"
)
maize_ftneg <- here("data", "maize_ftneg.sqlite")
con_mftneg <- dbConnect(RSQLite::SQLite(), maize_ftneg)

#maize FTpos

get_dynlib_zenodo(
  doi = ,
  file = "maize_ftmspos.sqlite",
  destdir = "data"
)
maize_ftpos <- here("data", "maize_ftpos.sqlite")
con_mftpos <- dbConnect(RSQLite::SQLite(), maize_ftpos)


# maize_QTOFneg
get_dynlib_zenodo(
  doi = ,
  file = "maize_qtofneg.sqlite",
  destdir = "data"
)
maize_qtofneg <- here("data", "maize_qtofneg.sqlite")
con_mqt <- dbConnect(RSQLite::SQLite(), maize_qtofneg)

#maize QTOFpos
get_dynlib_zenodo(
  doi = ,
  file = "maize_qtofpos.sqlite",
  destdir = "data"
)
maize_qtofpos <- here("data", "maize_qtofpos.sqlite")
con_mqtpos <- dbConnect(RSQLite::SQLite(), maize_qtofpos)

# flaxseed_FTneg
get_dynlib_zenodo(
  doi = ,
  file = "flax_ftms.sqlite",
  destdir = "data"
)
flax_ftneg <- here("data", "flax_ftneg.sqlite")
con_fft <- dbConnect(RSQLite::SQLite(), flax_ftneg)

# flaxseed_QTOFneg
get_dynlib_zenodo(
  doi = ,
  file = "flax_qtof.sqlite",
  destdir = "data"
)
flax_qtofneg <- here("data", "flax_qtofneg.sqlite")
con_fqt <- dbConnect(RSQLite::SQLite(), flax_qtofneg)

# HDG_FTneg 
get_dynlib_zenodo(
  doi = ,
  file = "hdg_ftms.sqlite",
  destdir = "data"
)
hdg_ftneg <- here("data", "hdg_ftneg.sqlite")
con_hdgft <- dbConnect(RSQLite::SQLite(), hdg_ftneg)

```

To create the target library for spectral matching, the maize FTneg DynLib database is merged with the HDG_FTneg DynLib database.

```{r}
target_DL_FTneg <- here("data", "FTneg_maize_hdg.sqlite")

#convert_add_rt is False because both databases are already in minutes so no need for conversion
merge_databases(main_db  = maize_ftneg, add_db = hdg_ftneg,
                output_db = target_DL_FTneg, convert_add_rt = FALSE)

con_merged <- dbConnect(RSQLite::SQLite(), target_DL_FTneg)

```

# Batch spectral similarity searching against a reference DynLib database

We start by MS2-MS2 similarity search of flax FTneg against maize+HDG FTneg (target_DL_FTneg)

## FTMS spectral similarity (flax vs maize_hdg)

First we convert both sql databases to a spectra object:

```{r}
sp_maize_hdg <- sql_to_spectra(target_DL_FTneg)
sp_ftflax <- sql_to_spectra(flax_ftneg)
```

A DynLib database can store different types of spectra for each compound. The spectrum types available in a given database can be inspected with `available_spectrum_types()`:

```{r}
available_spectrum_types(con_merged)
available_spectrum_types(con_fft) 
available_spectrum_types(con_qtofneg) 
```

Definitions of the different spectrum types are provided in the documentation of MultMatchCID_Spectra():

```{r}
?MultMatchCID_Spectra
```

We then use `similarity_RDynLib_match()` to perform spectral similarity searching between the query (new flaxseed experiment) and target DynLib spectra (maize and HDG). The fragmentation level and spectrum type can be specified independently for the query and target using the `msLevel_query`, `msLevel_target`, `spectrum_type_query`, and `spectrum_type_target` arguments. Other important arguments control the matching criteria. `fragments_resolution` determines the type of spectral similarity scoring calculation used. In this example, the fragmentation spectra in both the query and the target databases were acquired at unit resolution in the ion trap of a hybrid FT mass spectrometer. Therefore, we set `fragments_resolution = "unit.resolution"`. 

When `requirePrecursor = TRUE`, the `ppm` and `tolerance` arguments define the precursor m/z matching criteria that are applied before the spectral similarity calculation. The two tolerances are combined, such that precursor masses are considered matching when their absolute m/z difference is within tolerance + ppm(m/z). These parameters can therefore be adjusted according to the mass accuracy of the precursor ions being compared. For example, in our FT-MSⁿ data, MS2 precursor masses are measured at high resolution, whereas the precursor masses recorded for MS3 and MS4 spectra have unit resolution. Consequently, when comparing an MS3 flax spectrum with an MS2 spectrum from the maize-HDG DynLib, a larger absolute tolerance can be used to accommodate the lower mass accuracy of the MS3 precursor.

The available options and the corresponding scoring procedures are described in the documentation of `similarity_RDynLib_match()`.

```{r}
?similarity_RDynLib_match()
```

As an example of a comparison involving different MS levels, we first search the MS3 flax spectra against the MS2 spectra from the maize-HDG DynLib. Because the precursor m/z values associated with the MS3 spectra are recorded at unit resolution, we use a larger absolute precursor mass tolerance.

```{r}
ftms_matches3_T <- similarity_RDynLib_match(
  sp_ftflax,
  sp_maize_hdg,
  polarity_query = 0,
  polarity_target = 0,
  msLevel_query = 3,
  msLevel_target = 2,
  spectrum_type_query = "individual",
  spectrum_type_target = "individual",
  requirePrecursor = TRUE,
  tolerance = 0.5,
  ppm = 5,
  threshold = 0.7,
  fragments_resolution = "unit.resolution"
)
```

We have a quick look at the MS3-MS2 matches and order them according to decreasing similarity score and number of matched peaks.

```{r}
ftms_matches3_T |>
  dplyr::arrange(
    dplyr::desc(score),
    dplyr::desc(matched_peaks_count)
  )
```

This search illustrates how the precursor mass tolerance can be adapted when spectra from different MS levels are compared. For the remainder of this workflow, however, we focus on matching MS2 spectra from the flax experiment against MS2 spectra in the maize-HDG DynLib. Since the precursor m/z values of these spectra were measured at high resolution, a stricter precursor mass tolerance can be applied.

```{r}
ftms_matches2_T <- similarity_RDynLib_match(
  sp_ftflax,
  sp_maize_hdg,
  polarity_query = 0,
  polarity_target = 0,
  msLevel_query = 2,
  msLevel_target = 2,
  spectrum_type_query = "individual",
  spectrum_type_target = "individual",
  requirePrecursor = TRUE,
  tolerance = 0,
  ppm = 5,
  threshold = 0.7,
  fragments_resolution = "unit.resolution"
 )
```

We have a quick look at the found matches. We order them according to decreasing similarity scores based on the number of matched peaks.

```{r}
names(ftms_matches2_T)
nrow(ftms_matches2_T)
o <- order(ftms_matches2_T$matched_peaks_count, decreasing =T)
print(ftms_matches2_T[o,c("query_compound_id", "query_spectrum_id", "msLevel","precursorMz","target_name","target_msLevel","target_precursorMz","target_compound_id","score", "matched_peaks_count")])
 
```

We can save the spectral matching results as .csv file for later consultation. To do so, we select the relevant columns from the similarity search output and optionally add metadata, such as the query date, the databases used, and the matching parameters. Recording this information allows us to subsequently combine results from different similarity searches, potentially involving different databases, spectrum types, MS levels, or matching parameters. This facilitates the comparison, interpretation, and integration of results across multiple searches.

```{r}
# Select the variable of interest to save
(matches_table <- (ftms_matches2_T[o,c("query_compound_id", "query_compound_accession","query_spectrum_id", "msLevel","precursorMz","target_name","target_msLevel","target_precursorMz","target_compound_id","target_compound_accession","target_spectrum_id","score", "matched_peaks_count")]))

# Extend the table with metadata about the Query (date of query, query database, target databases)

# Retrieve the experiment accession
experiment_accession <- DBI::dbGetQuery(con_fft, "SELECT experiment_accession FROM experiment")

# Assemble the table
matches_table_assembled <- data.frame(experiment = experiment_accession, queryDate = "2009/09/30", queryDB = "flax_FTneg2023", targetDB ="maize_hdg_QTOFneg", matches_table)

# Export to a .csv file
write.csv(matches_table_assembled, here("data","matches_ftms_flaxseed_same_energy_vs_maizehdg_individual.csv"), row.names = FALSE)

```

Six query spectra yielded a perfect similarity score of 1, with more than five matched spectral peaks (matching m/z of fragments or neutral losses). The spectral matches, or a selected subset, can be stored in dedicated columns of the `msms_spectrum` table of the DynLib SQL database.

However, storing new matching results will overwrite any previously stored results for the corresponding query spectra. We are curious to know whether matching results already exist for the six spectra identified above.

Using their `query_spectrum_id` values, and a simple SQL query through the existing connection `con_ftt`, we retrieve the corresponding records from the `msms_spectrum` table of the `flax_ftneg` database. We then display only the columns containing the `spectrum identifiers`spectrum_id\` and spectral matching information.

```{r}
# Retrieve the query spectrum IDs of the six best matches
query_ids <- ftms_matches2_T$query_spectrum_id[o][1:6]

# Retrieve their existing records from the DynLib database
existing_matches <- DBI::dbGetQuery(
  con_fft,
  paste(
    "SELECT * FROM msms_spectrum WHERE spectrum_id IN (", paste(query_ids, collapse = ", "),")"
  )
)

existing_matches[,c("spectrum_id","matches","match_name","match_compound_id","match_precursor_mz","match_score","match_common","match_mslevel","match_spectrum_type")]
```

If we want to store store the best six matches in the flax ftneg SQL database (and overwrite existing matches if present):

```{r}
# Make a selection of the spectral similarity output table
ftms_matches_sel <- ftms_matches2_T[o,][1:6,]

# Add the selection of matches to the msms_spectrum table of the sql database
add_spectral_matches(con_fft, ftms_matches_sel)

# Verify the addition, printing only relevant columns
DBI::dbGetQuery(con_fft,
  paste(
    "Select spectrum_id, ms_level, compound_id, precursor_mz,
                     spectrum_type, matches, match_mslevel, match_common,
                     match_precursor_mz, match_compound_id, match_spectrum_type
           from msms_spectrum where spectrum_id IN (", paste(query_ids, collapse = ", "),")"
  )
)


```

The add_spectral_matches()\` function generates a summary of the spectral match, consisting of the name of the matched compound, prefixed with the number of matched peaks and the similarity score, and stores this value in the `matches` variable of the `msms_table`.

TODO change this column name to `match`.

You may want to have a look:

```{r}
DBI::dbGetQuery(con_fft,
  paste(
    "Select compound_id, matches from msms_spectrum where spectrum_id IN (", paste(query_ids, collapse = ", "),")"
  )
)
```

## QTOF spectral similarity (flax vs maize_hdg)

We can apply the same batch similarity matching procedure to QTOF data. We first need to convert the corresponding query and target sql databases to `Spectra` objects:

```{r}
sp_qtmaize <- sql_to_spectra(maize_qtofneg)
sp_qtflax <- sql_to_spectra(flax_qtofneg)
```

Next we extract the available spectrum types in both databases:

```{r}
available_spectrum_types(con_mqt) 
```

```{r}
available_spectrum_types(con_fqt) 
```

Then we call the similarity_RDynLib_match() function :

```{r}
qtof_matches2_T <- similarity_RDynLib_match(
  sp_qtflax,
  sp_qtmaize,
  polarity_query = 0,
  polarity_target = 0,
  msLevel_query = 2,
  msLevel_target = 2,
  spectrum_type_query = "same_energy",
  spectrum_type_target = "individual",
  requirePrecursor = TRUE,
  tolerance = 0,
  ppm = 20,
  threshold = 0.7,
  fragments_resolution = "high.resolution"
 )
```

Show the matches, and save them to a .csv file:

```{r}
o <- order(qtof_matches2_T$matched_peaks_count, decreasing =T)
print(qtof_matches2_T[o,c("query_compound_id", "query_spectrum_id", "msLevel","precursorMz","target_name","target_msLevel","target_precursorMz","target_compound_id","score", "matched_peaks_count")])

# We can save the table as .csv file for later consultation
# Select the variable of interest to save
(matches_table <- (qtof_matches2_T[o,c("query_compound_id", "query_compound_accession","query_spectrum_id", "msLevel","precursorMz","target_name","target_msLevel","target_precursorMz","target_compound_id","target_compound_accession","target_spectrum_id","score", "matched_peaks_count")]))

# Extend the table with metadata about the Query (date of query, query database, target databases)

# Retrieve the experiment accession
experiment_accession <- DBI::dbGetQuery(con_fqt, "SELECT experiment_accession FROM experiment")

# Assemble the table
matches_table_assembled <- data.frame(experiment = experiment_accession, queryDate = "2009/09/30", queryDB = "flax_QTOFneg2023", targetDB ="maize_hdg_QTOFneg", matches_table)

# Export to a .csv file
write.csv(matches_table_assembled, here("data","matches_qtof_flaxseed_same_energy_vs_maizehdg_individual.csv"), row.names = FALSE)

```

We can store the spectral matches in the `msms_spectrum` table of the query SQL database:

```{r}
# First check if there are spectral matches stored in the database
# Retrieve the query spectrum IDs
query_ids <- qtof_matches2_T$query_spectrum_id

# Retrieve their existing records from the DynLib database
existing_matches <- DBI::dbGetQuery(
  con_fqt,
  paste(
    "SELECT * FROM msms_spectrum WHERE spectrum_id IN (", paste(query_ids, collapse = ", "),")"
  )
)

existing_matches[,c("spectrum_id","matches","match_name","match_compound_id","match_precursor_mz","match_score","match_common","match_mslevel","match_spectrum_type")]

# If OK to overwrite, proceed:
add_spectral_matches(con_fqt, qtof_matches2_T)
  
# Verify the addition of the matching data in the database
dbGetQuery(con_fqt,
  paste(
    "Select spectrum_id, ms_level, compound_id, precursor_mz,
                     spectrum_type, matches, match_mslevel, match_common,
                     match_precursor_mz, match_compound_id, match_spectrum_type
           from msms_spectrum where spectrum_id IN (", paste(query_ids, collapse = ", "),")"
  )
)

```


# Batch spectral similarity searching against a public spectral database

In this section, spectra from the flax seed DynLib database are **exported in MGF format**, enabling their use with external **online tools** and **public spectral databases** for spectral similarity searching and further metabolite annotation.

A DynLib database can store different types of spectra for each compound; see the documentation of `MultMatchCID_Spectra()` for definitions of the different spectrum types. For multilevel fragmentation trees, such as those acquired with ion trap instruments, spectra can additionally be distinguished by their **MSⁿ level**. When exporting spectra to MGF format, RDynLib allows the user to select the **spectrum type** and, where applicable, the **fragmentation level** to be exported. In a multi-experiment DynLib database, the export can also be restricted to spectra from selected experiments.

The following example illustrates the selection of an experiment, spectrum type, and fragmentation level for MGF export. A DynLib database can store different types of spectra for each compound. For multilevel fragmentation trees, such as those acquired with ion trap instruments, spectra can additionally be distinguished by their MSⁿ level. When exporting spectra to MGF format, RDynLib allows the user to **select the spectrum type** and, where applicable, the **fragmentation level** to be exported. In a multi-experiment DynLib database, the export can be restricted to spectra from **selected experiments**.

## The following example illustrates the selection of an experiment, spectrum type, and fragmentation level for MGF export.
TODO


# Single spectrum spectral similarity search against DynLib
We start by the CID spectral matching of 101280 compound from ftms negative database, To do that we need to convert tha database to a spectra object:

```{r}
dy_sps <- sql_to_spectra(ftneg_path)
```

```{r}
cmp <- MultMatchCID_Spectra(
  dy_sps,
  101280,
   spectrum_type_query = "individual",
  spectrum_type_target = "individual",
   ms_level_query = 2,
   ms_level_target = 2,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0,
  ppm = 10,
  minIons = ,
    fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
head(cmp, 5)
```

Then we update the cspp table:

```{r}
#| message: false
#| results: "hide"
#| warning: false
cspp.tot_SQL(ftneg_path, 266, mzerr = 0.01,
                         cspp = cspp_ft, peakwidth = 0.1, IntThres  = 0, mz_tol = 0.5,
                         dot_thresh = 0, spectrum_type = "individual",  
                         min_common = 0, charge_file = charge_file)
```

Update the gnps table:

```{r}
#| message: false
#| results: "hide"
#| warning: false
conv.GNPS_SQL(sql_path= ftneg_path, expid = 266, peakwidth = 0.2, mzerr = 0.015,
              min = 5, adduct = 60.0211, thr1 = 2, thr2 = 0.1, thr3 = 0.3,
              spectrum_type = "all_energies", IntThres = 0)
```

Next we visualize the CSPP and GNPS network of 101280 compound from ftms neg:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 20
#| fig-height: 15
Overall.net_SQL(sql_path = ftneg_path, exp.id = 266, dbkey = 101280, min = 0,
                nr_of_seq = 3, thr1 = 0.1, thr2 = 0.2, thr3 = 0.3)
```

Now we show the ion trap based analysis of 100988 compound:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSnplot_SQL(ftneg_path, 100988, thr_int = 100)
```

```{r}
#| fig-width: 14
#| fig-height: 4
MSnplot_SQL(ftneg_path, 101155, thr_int = 100)
```

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSnplot_SQL(ftneg_path, 101283, thr_int = 100)
```


# Compound annotation in RDynLib

Based on the spectral matches, compound names can be automatically added to the `ms_compound` table using the `**adding_Names_SQL()**` function. The automatically generated name combines the name of the matched compound with the spectral similarity score and is prefixed with `!!` to distinguish it from manually assigned structural annotations. No confidence level is assigned at this stage.

Each automatic annotation also creates an **annotation event** in the `annotation_history` table. The event records information supporting the match, including the spectral similarity score and the experiment accession of the matched compound. While the name stored in `ms_compound` can subsequently be replaced as the compound is further characterized, the annotation event is retained in the annotation history, providing a persistent record of the spectral match and the progression of the compound annotation.

```{r}
adding_Names_SQL(con_fft,  spectrum_type = "individual", msLevel = 2, compound_id = NULL,
    overwrite = FALSE )
```

Adding names in ftms

```{r}
adding_Names_SQL(con_fft,  spectrum_type = "individual", msLevel = 2, compound_id = NULL,
    overwrite = FALSE )
```

#Qtof flax analysis

Import the database and required files:

```{r}
qt_path <- here("data", "DYL_FLAX_QTOFneg.sqlite")
con_qt <- dbConnect(RSQLite::SQLite(), qt_path )
neutral_loss_candid <- here("data", "neutral_loss_candid.txt")
neutloss <- read.table(
  neutral_loss_candid,
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

product_ion_candid <- here("data", "product_ion_candid_2.txt")
prdion <- read.table(
  product_ion_candid,
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

```

The aligned compound in qtof data that corresponds to 100988 compound from ftms data:

```{r}
alignment_file <- here("data", "assoc_flax_63.txt")

getAlignedCompids <- getAlignedCompids_SQL(
compound_id = 100988,
  database_name = "DYL_FLAX_FTMSneg.sqlite",
  alignment_file = alignment_file
)
getAlignedCompids
```

The 45890 is the qtof compound_id that was aligned with 100988 ftms compound_id.

The msms_spectrum table linked to that compound_id to select the spectrum_id:

```{r}
list_spectrum <- get_spectrum_id(qt_path,45890)
list_spectrum
```

We chose the 46001 spectrum_id to show the qtof based neutral loss analysis:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSMSspecplot_SQL(qt_path, 46001, 
                 prdion = prdion, neutloss = neutloss, err = NULL, 
                 minum = NULL, nr_col = NULL,   nr_col2 = NULL)
```

Next, we call the function to export the sql database to an mgf format that could be used for further analysis and comparisons with public databases available on public platforms like: GNPS... We start by converting ftms data

```{r}
#| message: false
#| results: "hide"
#| warning: false
ft_assembled <- here("data", "ft_assembled.mgf")
sql_to_mgf(ftneg_path, expid = 266, spectrum_type = "pseudo_MS2", 
           output_file = ft_assembled) 
```

Then, we convert qtof data:

```{r}
#| message: false
#| results: "hide"
#| warning: false
qt_assembled <- here("data", "qt_assembled.mgf")
sql_to_mgf(qt_path, expid = 52, spectrum_type = "same_energy", 
           output_file = qt_assembled) 
```

#Alignment RDinlib offres tools for aliging two expriment acquired from th same/diffrent platforms with the same/different ionization modes, as well as two different experiments stored in the same database. The aligned **compound_ids** are then stored in a given .txt file.

We start by aligning the flax_ftneg flax ftms negative data against the flax qtof negative data:

```{r}
alignment_output <- here("data", "flax_qtn_ftn_align.txt")

Assoc_SQL_flax <-Aligning_FT_QTOF_SQL(
  FT_path = flax_ftneg,
  QTOF_path = flax_qtofneg,
  FT_expnr =  266,
  QTOF_expnr = 52,
  Assoc = alignment_output,
  startpoint = 1,
  save_assoc = TRUE,
  spectrum_type_FT = "pseudo_MS2", 
  spectrum_type_QTOF = "all_energies",
  minIon = 0.4,
  minPeaks = 0.4,
  lc.err = 1,
  err = 0.02
)
nrow(Assoc_SQL_flax)
```

The following function extracts the aligned compound of a given compound_id:

```{r}
getAlignedCompids_SQL(
  compound_id = 101085,
  database_name = "flax_ftneg.sqlite",
  alignment_file = alignment_output
)
```

We can also print the corresponding row from the ms_compound table:

```{r}
cmp <- get_compound(qt_path = flax_ftneg, compound_id = 101085)
cmp
```

Next we use the function to align compounds from the same instrument and with different polarities, as a use case here we perform this alignment between maize ftms positive and maize ftms positive:

```{r}
ft_maize_alignment <- Aligning_FTn_FTp_SQL(FTn_path = maize_ftneg,
  FTp_path = maize_ftpos,
  FTn_expnr =  192,
  FTp_expnr = 56,
  Assoc = NULL,
  startpoint = 1,
  save_assoc = FALSE,
  spectrum_type_FTn = "individual", 
  spectrum_type_FTp = NULL,
  minIon = 0.4,
  minPeaks = 0.4,
  lc.err = 1,
  err = 0.02)
nrow(ft_maize_alignment)
```

Now we align two experiment stored in the same database(acquired from the same instrument and polarity), here we test the function on two experiments from the maize ftms negative:

```{r}
align_2_exp <- Aligning_Exp_SQL(
    db_path = maize_ftneg,
    ref_expnr = 195,
    target_expnr = 196,
    Assoc = NULL,
    err = 0.01,
    t.ini = 5,
    lc.err = 1,
    rng = 2,
    thr1 = 1,
    thr2 = 0.9,
    IntThres = 100,
    startpoint = 1,
    tol = 0.01,
    save_assoc = TRUE,
    spectrum_type_ref = "individual", 
    spectrum_type_target = "individual"
)
nrow(align_2_exp)
```

After testing the different types of the alignement functions in RDynlib, now we can transfer the names of the aligned compounds between databases: We test the function on the flax ftms data

```{r}
dir <- here("data")
SubDBNameTransfer_SQL(sqlite_dir = dir,  assoc_path = alignment_output )
```

Now we enter the characterization of the compounds found to the database., Using EnterName fucntion we are able to edit compound names and add additional structural information, including molecular formula, adduct type, and SMILES notation. Together, these functions provide a framework for progressive annotation of DynLib databases, ranging from automated transfer of tentative annotations to manual curation of fully characterized compounds:

```{r}
EnterName_SQL(maize_qtofpos, CoMPID = 1144)
```

```{r}
#EnterName_SQL(ft_path, CoMPID = 100988)
```
