---
title: "RDynLib workflow for metabolite structural annotation"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe"
---

# Overview

This workflow illustrates how the **RDynLib infrastructure supports metabolite structural annotation**, from spectral data exploration and comparison to the recording of compound annotations and their provenance in DynLib databases.

A flax seed DynLib database, generated from a new LC-MS experiment containing compounds to be structurally annotated, serves as the **query database** throughout the workflow. It is compared with a maize DynLib database containing previously characterized metabolites and their associated spectra, which serves as the **reference database**.

The workflow progresses from general **spectral similarity searching** against a reference DynLib database (the maize database in this example) and the **export of spectral data for further analysis with external databases and tools**, to detailed comparison and **visualization of individual spectra**, including **matching across different MSⁿ levels**. A key functionality of RDynLib is the **joint exploitation of complementary fragmentation information acquired on different mass spectrometry platforms**, such as ion trap and QTOF instruments, to support structural characterization. Biochemical and spectral relationships are explored through the construction and visualization of **CSPP and GNPS-like molecular networks**.

Finally, the resulting **compound annotation is written to the flaxseed DynLib database**. Along with the structural annotation itself, the evidence and reasoning underlying the characterization are recorded as an **annotation event**, preserving the provenance of the annotation and documenting the justification for the **assigned confidence level**.

# Workflow setup

## Loading required packages

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| 
library(devtools)
library(DBI)
library(RSQLite)
library(here)
library(CompoundDb)
library(MsBackendMgf)

# TODO: Replace with library(RDynLib) once the package is finalized.
load_all()
```

## Retrieving DynLib datasets from Zenodo

The DynLib SQLite databases used in this workflow are provided through Zenodo. Each database contains the features detected in one or more LC-MS experiments together with their associated mass spectral data and experimental metadata and compound annotations. Separate databases are used for each mass spectrometry platform and ionization mode:

-   **FTneg / FTpos**: LTQ-Orbitrap FT-MS data acquired in negative / positive ionization mode, including MSⁿ spectra
-   **QTOFneg / QTOFpos**: QTOF-MS data acquired in negative / positive ionization mode, including MS/MS spectra

```{r}
# TODO: this chunk is for the working document and needs to be replaced by the Zenodo chunk when we will have deposited the databases.
#| message: false
#| results: "hide"
#| warning: false
ftneg_path <- here("data", "DYL_FLAX_herba_FTMSneg.sqlite")
con_ftneg <- dbConnect(RSQLite::SQLite(), ftneg_path)

# flax QTOFneg
qtofneg_path <- here("data", "DYL_FLAX_QTOFneg.sqlite")
con_qtofneg <- dbConnect(RSQLite::SQLite(), qtofneg_path)

# maize FTneg
maize_ftneg <- here("data", "maize_ftneg.sqlite")
filterDynLibDB(
  expid = 191:231,
  input_db = ftneg_path,
  output_db = maize_ftneg
)
con_mftneg <- dbConnect(RSQLite::SQLite(), maize_ftneg)

# maize QTOFneg
maize_qtofneg <- here("data", "maize_qtofneg.sqlite")
filterDynLibDB(
  expid = 38,
  input_db = qtofneg_path,
  output_db = maize_qtofneg
)
con_mqt <- dbConnect(RSQLite::SQLite(), maize_qtofneg)

# flax FTneg
flax_ftneg <- here("data", "flax_ftneg.sqlite")
filterDynLibDB(
  expid = 266,
  input_db = ftneg_path,
  output_db = here("data", "flax_ftneg.sqlite")
)
con_fft <- dbConnect(RSQLite::SQLite(), flax_ftneg)

# flax QTOFneg
flax_qtofneg <- here("data", "flax_qtofneg.sqlite")
filterDynLibDB(
  expid = 52,
  input_db = qtofneg_path,
  output_db = here("data", "flax_qtofneg.sqlite")
)
con_fqt <- dbConnect(RSQLite::SQLite(), flax_qtofneg)

# HDG FTneg 
hdg_ftneg <- here("data", "hdg_ftneg.sqlite")
filterDynLibDB(
  expid = 267,
  input_db = ftneg_path,
  output_db = here("data", "hdg_ftneg.sqlite"))
con_hdgft <- dbConnect(RSQLite::SQLite(), hdg_ftneg)

```

```{r}
# TODO: this chunk is for the working document and needs to be replaced by the Zenodo chunk when we will have deposited the databases.
#| message: false
#| results: "hide"
#| warning: false
# Import the data from zenodo and create the connections
# maize_FTneg
get_dynlib_zenodo(
  doi = ,
  file = "maize_ftms.sqlite",
  destdir = "data"
)
maize_ftneg <- here("data", "maize_ftms.sqlite")
con_mft <- dbConnect(RSQLite::SQLite(), maize_ftneg)

# maize_QTOFneg
get_dynlib_zenodo(
  doi = ,
  file = "maize_qtof.sqlite",
  destdir = "data"
)
maize_qtofneg <- here("data", "maize_qtof.sqlite")
con_mqt <- dbConnect(RSQLite::SQLite(), maize_qtofneg)

# flaxseed_FTneg
get_dynlib_zenodo(
  doi = ,
  file = "flax_ftms.sqlite",
  destdir = "data"
)
flax_ftneg <- here("data", "flax_ftms.sqlite")
con_fft <- dbConnect(RSQLite::SQLite(), flax_ftneg)

# flaxseed_QTOFneg
get_dynlib_zenodo(
  doi = ,
  file = "flax_qtof.sqlite",
  destdir = "data"
)
flax_qtofneg <- here("data", "flax_qtof.sqlite")
con_fqt <- dbConnect(RSQLite::SQLite(), flax_qtofneg)

# HDG_FTneg 
get_dynlib_zenodo(
  doi = ,
  file = "hdg_ftms.sqlite",
  destdir = "data"
)
hdg_ftneg <- here("data", "hdg_ftms.sqlite")
con_hdgft <- dbConnect(RSQLite::SQLite(), hdg_ftneg)

#Ahlam : we don't have hdg in qtof
```

To create the target library for spectral matching, the maize FTneg DynLib database is merged with the HDG_FTneg DynLib database.

```{r}
target_DL_FTneg <- here("data", "FTneg_maize_hdg.sqlite")

#convert_add_rt is False because both databases are already in minutes so no need for conversion
merge_databases(main_db  = maize_ftneg, add_db = hdg_ftneg,
                output_db = target_DL_FTneg, convert_add_rt = FALSE)

con_merged <- dbConnect(RSQLite::SQLite(), target_DL_FTneg)

```

# Spectral similarity searching against a reference DynLib database

We start by MS2-MS2 similarity search of flax FTneg against maize+HDG FTneg (target_DL_FTneg)

## FTMS spectral similarity (maize_hdg vs flax)

First we convert both sql databases to a spectra object:

```{r}
sp_maize_hdg <- sql_to_spectra(target_DL_FTneg)
sp_ftflax <- sql_to_spectra(flax_ftneg)
```

A DynLib database can store different types of spectra for each compound. The spectrum types available in a given database can be inspected with `available_spectrum_types()`:

```{r}
available_spectrum_types(con_merged)
available_spectrum_types(con_fft) 
```


Definitions of the different spectrum types are provided in the documentation of MultMatchCID_Spectra():

```{r}
?MultMatchCID_Spectra
```

We then use `similarity_RDynLib_match()` to perform spectral similarity searching between the query (new flaxseed experiment) and target DynLib spectra (maize and HDG). The fragmentation level and spectrum type can be specified independently for the query and target using the `msLevel_query`, `msLevel_target`, `spectrum_type_query`, and `spectrum_type_target` arguments. Other important arguments control the matching criteria. `fragments_resolution` determines the type of spectral similarity scoring calculation used. The available options and the corresponding scoring procedures are described in the documentation of `similarity_RDynLib_match()`.

```{r}
?similarity_RDynLib_match()
```

```{r}
ftms_matches2_T <- similarity_RDynLib_match(
  sp_ftflax,
  sp_maize_hdg,
  polarity_query = 0,
  polarity_target = 0,
  msLevel_query = 2,
  msLevel_target = 2,
  spectrum_type_query = "not_assembled",
  spectrum_type_target = "not_assembled",
  requirePrecursor = TRUE,
  tolerance = 0,
  ppm = 5,
  threshold = 0.7,
  fragments_resolution = "unit.resolution"
 )
```

Show the number of the found matches

```{r}
print(ftms_matches2_T)
nrow(ftms_matches2_T) 
```

Store the first 20 matches based on the matched_peaks_count:
```{r}
ft_20_matches <- ftms_matches2_T %>%
  arrange(desc(matched_peaks_count)) %>%
  slice_head(n = 20)
```

The spectral matches can be added to the `msms_spectrum` table of the sql database.

```{r}
add_spectral_matches(con_fft, ftms_matches2_T)

# Verify the addition

# TODO : print only selected relevant columns

dbGetQuery(con_fft, "Select spectrum_id, ms_level, compound_id, precursor_mz,
                     spectrum_type, matches, match_mslevel, match_common,
                     match_precursor_mz, match_compound_id, match_spectrum_type,
                     match_spectrum_id
           from msms_spectrum where match_name IS not null LIMIT 1")

```

Based on the spectral matches, compound names can be automatically added to the `ms_compound` table using the `**TODO()**` function. The automatically generated name combines the name of the matched compound with the spectral similarity score and is prefixed with `!!` to distinguish it from manually assigned structural annotations. No confidence level is assigned at this stage.

Each automatic annotation also creates an **annotation event** in the `annotation_history` table. The event records information supporting the match, including the spectral similarity score and the experiment accession of the matched compound. While the name stored in `ms_compound` can subsequently be replaced as the compound is further characterized, the annotation event is retained in the annotation history, providing a persistent record of the spectral match and the progression of the compound annotation.
```{r}
adding_Names_SQL(con_fft,  spectrum_type = "not_assembled", msLevel = 2, compound_id = NULL,
    overwrite = FALSE )
```

## QTOF spectral similarity (maize vs flax)

First we convert both sql databases to a spectra object:

```{r}
sp_qtmaize <- sql_to_spectra(maize_qtofneg)
sp_qtflax <- sql_to_spectra(flax_qtofneg)
```

Next we extract the available spectrum types in both databases:

```{r}
available_spectrum_types(con_mqt) 
```

```{r}
available_spectrum_types(con_fqt) 
```

Then we call the similarity_RDynLib_match() function :

```{r}
qtof_matches2_T <- similarity_RDynLib_match(
  sp_qtflax,
  sp_qtmaize,
  polarity_query = 0,
  polarity_target = 0,
  msLevel_query = 2,
  msLevel_target = 2,
  spectrum_type_query = "intersect_single_energy",
  spectrum_type_target = "not_assembled",
  requirePrecursor = TRUE,
  tolerance = 0,
  ppm = 5,
  threshold = 0.7,
  fragments_resolution = "high.resolution"
 )
```


Show the matches:
```{r}
print(qtof_matches2_T)
nrow(qtof_matches2_T) 
```

Store the first 20 matches based on the matched_peaks_count:
```{r}
qt_20_matches <- qtof_matches2_T %>%
  arrange(desc(matched_peaks_count)) %>%
  slice_head(n = 20)
```

The spectral matches can be added to the `msms_spectrum` table of the sql database.

```{r}
add_spectral_matches(con_fqt, qtof_matches2_T)
```

In case of matches, verify if they are well added to the database:

```{r}
dbGetQuery(con_fqt, "Select spectrum_id, ms_level, compound_id, precursor_mz,
                     spectrum_type, matches, match_mslevel, match_common,
                     match_precursor_mz, match_compound_id, match_spectrum_type,
                     match_spectrum_id
           from msms_spectrum where match_name IS not null LIMIT 1")
```

Adding names as ftms
```{r}
adding_Names_SQL(con_fft,  spectrum_type = "not_assembled", msLevel = 2, compound_id = NULL,
    overwrite = FALSE )
```

# Spectral similarity searching against a public spectral database

In this section, spectra from the flax seed DynLib database are **exported in MGF format**, enabling their use with external **online tools** and **public spectral databases** for spectral similarity searching and further metabolite annotation.

A DynLib database can store different types of spectra for each compound; see the documentation of `MultMatchCID_Spectra()` for definitions of the different spectrum types. For multilevel fragmentation trees, such as those acquired with ion trap instruments, spectra can additionally be distinguished by their **MSⁿ level**. When exporting spectra to MGF format, RDynLib allows the user to select the **spectrum type** and, where applicable, the **fragmentation level** to be exported. In a multi-experiment DynLib database, the export can also be restricted to spectra from selected experiments.

The following example illustrates the selection of an experiment, spectrum type, and fragmentation level for MGF export. A DynLib database can store different types of spectra for each compound. For multilevel fragmentation trees, such as those acquired with ion trap instruments, spectra can additionally be distinguished by their MSⁿ level. When exporting spectra to MGF format, RDynLib allows the user to **select the spectrum type** and, where applicable, the **fragmentation level** to be exported. In a multi-experiment DynLib database, the export can be restricted to spectra from **selected experiments**.

# The following example illustrates the selection of an experiment, spectrum type, and fragmentation level for MGF export.

We start by the CID spectral matching of 101280 compound from ftms negative database, To do that we need to convert tha database to a spectra object:

```{r}
dy_sps <- sql_to_spectra(ftneg_path)
```

```{r}
cmp <- MultMatchCID_Spectra(
  dy_sps,
  101280,
   spectrum_type_query = "not_assembled",
  spectrum_type_target = "not_assembled",
   ms_level_query = 2,
   ms_level_target = 2,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0,
  ppm = 10,
  minIons = ,
    fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
head(cmp, 5)
```

Then we update the cspp table:

```{r}
#| message: false
#| results: "hide"
#| warning: false
cspp.tot_SQL(ftneg_path, 266, mzerr = 0.01,
                         cspp = cspp_ft, peakwidth = 0.1, IntThres  = 0, mz_tol = 0.5,
                         dot_thresh = 0, spectrum_type = "assembled",  
                         min_common = 0, charge_file = charge_file)
```

Update the gnps table:

```{r}
#| message: false
#| results: "hide"
#| warning: false
conv.GNPS_SQL(sql_path= ftneg_path, expid = 266, peakwidth = 0.2, mzerr = 0.015,
              min = 5, adduct = 60.0211, thr1 = 2, thr2 = 0.1, thr3 = 0.3,
              spectrum_type = "not_assembled", IntThres = 0)
```

Next we visualize the CSPP and GNPS network of 101280 compound from ftms neg:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 20
#| fig-height: 15
Overall.net_SQL(sql_path = ftneg_path, exp.id = 266, dbkey = 101280, min = 0,
                nr_of_seq = 3, thr1 = 0.1, thr2 = 0.2, thr3 = 0.3)
```

Now we show the ion trap based analysis of 100988 compound:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSnplot_SQL(ftneg_path, 100988, thr_int = 100)
```

```{r}
#| fig-width: 14
#| fig-height: 4
MSnplot_SQL(ftneg_path, 101155, thr_int = 100)
```

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSnplot_SQL(ftneg_path, 101283, thr_int = 100)
```

#Qtof flax analysis

Import the database and required files:

```{r}
qt_path <- here("data", "DYL_FLAX_QTOFneg.sqlite")
con_qt <- dbConnect(RSQLite::SQLite(), qt_path )
neutral_loss_candid <- here("data", "neutral_loss_candid.txt")
neutloss <- read.table(
  neutral_loss_candid,
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

product_ion_candid <- here("data", "product_ion_candid_2.txt")
prdion <- read.table(
  product_ion_candid,
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

```

The aligned compound in qtof data that corresponds to 100988 compound from ftms data:

```{r}
alignment_file <- here("data", "assoc_flax_63.txt")

getAlignedCompids <- getAlignedCompids_SQL(
compound_id = 100988,
  database_name = "DYL_FLAX_FTMSneg.sqlite",
  alignment_file = alignment_file
)
getAlignedCompids
```

The 45890 is the qtof compound_id that was aligned with 100988 ftms compound_id.

The msms_spectrum table linked to that compound_id to select the spectrum_id:

```{r}
list_spectrum <- get_spectrum_id(45890)
list_spectrum
```

We chose the 46001 spectrum_id to show the qtof based neutral loss analysis:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSMSspecplot_SQL(qt_path, 46001, 
                 prdion = prdion, neutloss = neutloss, err = NULL, 
                 minum = NULL, nr_col = NULL,   nr_col2 = NULL)
```

Now we enter the characterization of the compounds found to the database:

```{r}
#EnterName_SQL(ft_path, CoMPID = 101143)
```

```{r}
#EnterName_SQL(ft_path, CoMPID = 100988)
```

Next, we call the function to export the sql database to an mgf format that could be used for further analysis and comparisons with public databases available on public platforms like: GNPS... We start by converting ftms data

```{r}
#| message: false
#| results: "hide"
#| warning: false
ft_assembled <- here("data", "ft_assembled.mgf")
sql_to_mgf(ftneg_path, expid = 266, spectrum_type = "assembled", 
           output_file = ft_assembled) 
```

Then, we convert qtof data:

```{r}
#| message: false
#| results: "hide"
#| warning: false
qt_assembled <- here("data", "qt_assembled.mgf")
sql_to_mgf(qt_path, expid = 52, spectrum_type = "intersect_single_energy", 
           output_file = qt_assembled) 
```
