---
title: "RDynLib workflow for metabolite structural annotation"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe"
---

# Overview

This workflow illustrates how the **RDynLib infrastructure supports metabolite structural annotation**, from spectral data exploration and comparison to the recording of compound annotations and their provenance in DynLib databases.

A flax seed DynLib database, generated from a new LC-MS experiment containing compounds to be structurally annotated, serves as the **query database** throughout the workflow. It is compared with a maize DynLib database containing previously characterized metabolites and their associated spectra, which serves as the **reference database**.

The workflow progresses from general **spectral similarity searching** against a reference DynLib database (the maize database in this example) and the **export of spectral data for further analysis with external databases and tools**, to detailed comparison and **visualization of individual spectra**, including **matching across different MSⁿ levels**. A key functionality of RDynLib is the **joint exploitation of complementary fragmentation information acquired on different mass spectrometry platforms**, such as ion trap and QTOF instruments, to support structural characterization. Biochemical and spectral relationships are explored through the construction and visualization of **CSPP and GNPS-like molecular networks**.

Finally, the resulting **compound annotation is written to the flaxseed DynLib database**. Along with the structural annotation itself, the evidence and reasoning underlying the characterization are recorded as an **annotation event**, preserving the provenance of the annotation and documenting the justification for the **assigned confidence level**.

# Workflow setup

## Loading required packages

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| 
library(devtools)
library(DBI)
library(RSQLite)
library(here)
library(CompoundDb)
library(MsBackendMgf)

# TODO: Replace with library(RDynLib) once the package is finalized.
load_all()
```

## Retrieving DynLib datasets from Zenodo

The DynLib SQLite databases used in this workflow are provided through Zenodo. Each database contains the features detected in one or more LC-MS experiments together with their associated mass spectral data and experimental metadata and compound annotations. Separate databases are used for each mass spectrometry platform and ionization mode:

-   **FTneg / FTpos**: LTQ-Orbitrap FT-MS data acquired in negative / positive ionization mode, including MSⁿ spectra
-   **QTOFneg / QTOFpos**: QTOF-MS data acquired in negative / positive ionization mode, including MS/MS spectra

```{r}
# TODO: this chunk is for the working document and needs to be replaced by the Zenodo chunk when we will have deposited the databases.
#| message: false
#| results: "hide"
#| warning: false
ft_path <- here("data", "DYL_FLAX_herba_FTMSneg.sqlite")
con_ft <- dbConnect(RSQLite::SQLite(), ft_path)

# TODO # Ahlam : 
# create also a path to the QTOFneg database, and by subsetting create the following databases:
#Qtof database
qt_path <- here("data", "DYL_FLAX_QTOFneg.sqlite")
con_qt <- dbConnect(RSQLite::SQLite(), qt_path)

# maize_FTneg
maize_ftms <- here("data", "maize_ftms.sqlite")
filterDynLibDB(
  expid = ,
  input_db = ft_path,
  output_db = maize_ftms
)
con_mft <- dbConnect(RSQLite::SQLite(), maize_ftms)

# maize_QTOFneg
maize_qtof <- here("data", "maize_qtof.sqlite")
filterDynLibDB(
  expid = ,
  input_db = ft_path,
  output_db = maize_qtof
)
con_mqt <- dbConnect(RSQLite::SQLite(), maize_qtof)

# flaxseed_FTneg
flax_ftms <- here("data", "flax_ftms.sqlite")
filterDynLibDB(
  expid = 266,
  input_db = ft_path,
  output_db = here("data", "flax_ftms.sqlite")
)
con_fft <- dbConnect(RSQLite::SQLite(), flax_ftms)
# flaxseed_QTOFneg
flax_qtof <- here("data", "flax_qtof.sqlite")
filterDynLibDB(
  expid = 52,
  input_db = input_db,
  output_db = here("data", "flax_qtof.sqlite")
)
con_fqt <- dbConnect(RSQLite::SQLite(), flax_qtof)

# HDG_FTneg 
hdg_ftms <- here("data", "hdg_ftms.sqlite")
filterDynLibDB(
  expid = 267,
  input_db = ft_path,
  output_db = here("data", "hdg_ftms.sqlite")
con_hdgft <- dbConnect(RSQLite::SQLite(), hdg_ftms)

#Ahlam : we don't have hdg in qtof
```

```{r}
# TODO: this chunk is for the working document and needs to be replaced by the Zenodo chunk when we will have deposited the databases.
#| message: false
#| results: "hide"
#| warning: false
ft_path <- here("data", "DYL_FLAX_herba_FTMSneg.sqlite")
con <- dbConnect(RSQLite::SQLite(), ft_path)

# TODO # Ahlam : 
# create also a path to the QTOFneg database, and by subsetting create the following databases:
# maize_FTneg
# maize_QTOFneg
# flaxseed_FTneg
# flaxseed_QTOFneg
# HDG_FTneg

```

To create the target library for spectral matching, the maize FTneg DynLib database is merged with the HDG_FTneg DynLib database.

```{r}
# TODO # Ahlam : 
#FTMS 
target_DL_FTneg <- here("data", "FT_maize_hdg.sqlite")

merge_databases(main_db  = maize_ftms, add_db = hdg_ftms,
                output_db = target_DL_FTneg)

con_merged <- dbConnect(RSQLite::SQLite(), target_DL_FTneg)

```

# Spectral similarity searching against a reference DynLib database

TODO: Ahlam: insert here the code for MS2-MS2 similarity search of flax FTneg against maize+HDG FTneg (target_DL_FTneg) similarity search of flax QTOFneg against maize QTOFneg (=target_DL_QTOFneg ) As optional code, add also Adding_names () function
 
Ahlam: for qtof similarity we don't have hdg in qtof so we will just calculate the similarity between qtof flax and qtof maize?

We start by MS2-MS2 similarity search of flax FTneg against maize+HDG FTneg
(target_DL_FTneg) 

#FTMS spectra similarity(maize_hdg vs flax)

First we convert both sql databases to a spectra object:
```{r}
sp_maize_hdg <- sql_to_spectra(target_DL_FTneg)
sp_ftflax <- sql_to_spectra(flax_ftms)
```

Next we extract the available spectrum types in both databases
```{r}
available_spectrum_types(con_merged)
```

```{r}
available_spectrum_types(con_fft) 
```

Then we call the similarity_RDynLib_match() function : 
```{r}
ftms_matches2_T <- similarity_RDynLib_match(
  sp_ftflax,
  sp_maize_hdg,
  polarity_query = 0,
  polarity_target = 0,
  msLevel_query = 2,
  msLevel_target = 2,
  spectrum_type_query = "not_assembled",
  spectrum_type_target = "not_assembled",
  requirePrecursor = TRUE,
  tolerance = 0,
  ppm = 5,
  threshold = 0.7,
  fragments_resolution = "unit.resolution"
 )
```

Show the number of the found matches
```{r}
print(ftms_matches2_T)
nrow(ftms_matches2_T) 
```

Adding names of the new matches to flax ftms sql database:
```{r}
add_spectral_matches(con_fft, ftms_matches2_T)
```

In case of matches, verify if they are well added to the database:
```{r}
dbGetQuery(con_fft, "Select * from msms_spectrum where match_name IS not null LIMIT 1")
```

#QTOF spectral similarity (maize vs flax)

First we convert both sql databases to a spectra object:
```{r}
sp_qtmaize <- sql_to_spectra(maize_qtof)
sp_qtflax <- sql_to_spectra(flax_qtof)
```

Next we extract the available spectrum types in both databases
```{r}
available_spectrum_types(con_mqt)
```

```{r}
available_spectrum_types(con_fqt) 
```

Then we call the similarity_RDynLib_match() function : 
```{r}
qtof_matches2_T <- similarity_RDynLib_match(
  sp_qtflax,
  sp_qtmaize,
  polarity_query = 0,
  polarity_target = 0,
  msLevel_query = 2,
  msLevel_target = 2,
  spectrum_type_query = "Intersect_single_energy",
  spectrum_type_target = "Intersect_single_energy",
  requirePrecursor = TRUE,
  tolerance = 0,
  ppm = 5,
  threshold = 0.7,
  fragments_resolution = "high.resolution"
 )
```

Show the number of the found matches
```{r}
print(qtof_matches2_T)
nrow(qtof_matches2_T) 
```

Adding names of the new matches to flax qtof sql database:
```{r}
add_spectral_matches(con_fqt, qtof_matches2_T)
```

In case of matches, verify if they are well added to the database:
```{r}
dbGetQuery(con_fqt, "Select * from msms_spectrum where match_name IS not null LIMIT 1")
```
=====================================================================

We start by the CID spectral matching of 101280 compound from ftms negative database, To do that we need to convert tha database to a spectra object:

```{r}
dy_sps <- sql_to_spectra(ft_path)
```

```{r}
cmp <- MultMatchCID_Spectra(
  dy_sps,
  101280,
   spectrum_type_query = "not_assembled",
  spectrum_type_target = "not_assembled",
   ms_level_query = 2,
   ms_level_target = 2,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0,
  ppm = 10,
  minIons = ,
    fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
head(cmp, 5)
```

Then we update the cspp table:

```{r}
#| message: false
#| results: "hide"
#| warning: false
cspp.tot_SQL(ft_path, 266, mzerr = 0.01,
                         cspp = cspp_ft, peakwidth = 0.1, IntThres  = 0, mz_tol = 0.5,
                         dot_thresh = 0, spectrum_type = "assembled",  
                         min_common = 0, charge_file = charge_file)
```

Update the gnps table:

```{r}
#| message: false
#| results: "hide"
#| warning: false
conv.GNPS_SQL(sql_path= ft_path, expid = 266, peakwidth = 0.2, mzerr = 0.015,
              min = 5, adduct = 60.0211, thr1 = 2, thr2 = 0.1, thr3 = 0.3,
              spectrum_type = "not_assembled", IntThres = 0)
```

Next we visualize the CSPP and GNPS network of 101280 compound from ftms neg:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 20
#| fig-height: 15
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 101280, min = 0,
                nr_of_seq = 3, thr1 = 0.1, thr2 = 0.2, thr3 = 0.3)
```

Now we show the ion trap based analysis of 100988 compound:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSnplot_SQL(ft_path, 100988, thr_int = 100)
```

```{r}
#| fig-width: 14
#| fig-height: 4
MSnplot_SQL(ft_path, 101155, thr_int = 100)
```

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSnplot_SQL(ft_path, 101283, thr_int = 100)
```

#Qtof flax analysis

Import the database and required files:

```{r}
qt_path <- here("data", "DYL_FLAX_QTOFneg.sqlite")
con_qt <- dbConnect(RSQLite::SQLite(), qt_path )
neutral_loss_candid <- here("data", "neutral_loss_candid.txt")
neutloss <- read.table(
  neutral_loss_candid,
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

product_ion_candid <- here("data", "product_ion_candid_2.txt")
prdion <- read.table(
  product_ion_candid,
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

```

The aligned compound in qtof data that corresponds to 100988 compound from ftms data:

```{r}
alignment_file <- here("data", "assoc_flax_63.txt")

getAlignedCompids <- getAlignedCompids_SQL(
compound_id = 100988,
  database_name = "DYL_FLAX_FTMSneg.sqlite",
  alignment_file = alignment_file
)
getAlignedCompids
```

The 45890 is the qtof compound_id that was aligned with 100988 ftms compound_id.

The msms_spectrum table linked to that compound_id to select the spectrum_id:

```{r}
list_spectrum <- get_spectrum_id(45890)
list_spectrum
```

We chose the 46001 spectrum_id to show the qtof based neutral loss analysis:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSMSspecplot_SQL(qt_path, 46001, 
                 prdion = prdion, neutloss = neutloss, err = NULL, 
                 minum = NULL, nr_col = NULL,   nr_col2 = NULL)
```

Now we enter the characterization of the compounds found to the database:

```{r}
#EnterName_SQL(ft_path, CoMPID = 101143)
```

```{r}
#EnterName_SQL(ft_path, CoMPID = 100988)
```

Next, we call the function to export the sql database to an mgf format that could be used for further analysis and comparisons with public databases available on public platforms like: GNPS... We start by converting ftms data

```{r}
#| message: false
#| results: "hide"
#| warning: false
ft_assembled <- here("data", "ft_assembled.mgf")
sql_to_mgf(ft_path, expid = 266, spectrum_type = "assembled", 
           output_file = ft_assembled) 
```

Then, we convert qtof data:

```{r}
#| message: false
#| results: "hide"
#| warning: false
qt_assembled <- here("data", "qt_assembled.mgf")
sql_to_mgf(qt_path, expid = 52, spectrum_type = "intersect_single_energy", 
           output_file = qt_assembled) 
```
