---
title: "workflow to test the similarity_RDynlib() function on qtof data"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This quarto document represents the workflow for calling the `similarity_RDynlib()` from the [similarity_functions.R](https://github.com/rebdau/RDynLib_SQL/blob/Ahlam/R/similarity_functions.R) file to calculate the similarity between QTOF flax data and the RDynLib QTOFdatabase.

# Required packages

```{r}
library(here)
library(Spectra)
library(CompoundDb)
library(MetaboAnnotation)
library(DBI)
library(RSQLite)
library(devtools)
library(dplyr)
library(stringr)
#` Uploading RDynLibSQL: Load all functions of the RDynlib directory
load_all()
```

Then, install the new CompoundDb version:

```{r}
BiocManager::install("RforMassSpectrometry/CompoundDb", ref = "jomain", force = TRUE)
```

## convert the "DyLib_QTOF_neg.sqlite" database to a spectra object

Next we convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object
```{r}

pth <- here("data", "DYL_FLAX_QTOFneg.sqlite")
con <- dbConnect(SQLite(), pth)
dbExecute(
  con,
  "UPDATE msms_spectrum
   SET spectrum_type = 'not_assembled'
   WHERE spectrum_type IS NULL"
)

qt_dy_sps <- sql_to_spectra(pth)
```

Here we load the QTOF of flax seed database (it's already a spectra object), you can load it from: [qtof_filtering.qmd](https://github.com/rebdau/RDynLib_qtof/blob/ahlam/qtof_filtering.qmd)

```{r}
load(here("data/database_QTOF_neg","qtof_flax_final.RData"))
qtof_neg <- qtof_flax_final
spectraData(qtof_neg)
```

Next we calculate the similarity score between the dynlib and flax seed QTOF negative integuments.(intersect_single_energy)

```{r}
qtof_matchesF <- similarity_RDynLib_match(
  qtof_neg,
  qt_dy_sps,
  polarity_query = 0,
  polarity_target = 0,
  fragments_resolution = "high.resolution",
  threshold = 0.8,
  ppm = 10,
  tolerance = 0,
  requirePrecursor = FALSE
)
```

print the similarity matrix and the number of rows (matches)

```{r}
print(qtof_matchesF)
nrow(qtof_matchesF)
```

Adding names of the new matches:

```{r}
pth <- here("data", "DYL_FLAX_QTOFneg.sqlite")
qt_con <- dbConnect(SQLite(), pth)
```


```{r}
add_spectral_matches(qt_con, qtof_matchesF)
```

```{r}
dbGetQuery(qt_con, "Select * from msms_spectrum where match_name IS not null LIMIT 1")
```

look for a given name
```{r}
qtof_matchesF %>%
  filter(str_detect(target_name, regex("trans-aconitic acid AN KM", ignore_case = TRUE)))
```


Select the first 10 matches with the highest common peaks
```{r}
best_hits2 <- qtof_matchesF %>%
    arrange(desc(matched_peaks_count)) %>%   
    slice_head(n = 10)                

best_hits
```

```{r}
best_hits2 %>%
    select(
        
        precursorMz,
        target_precursorMz,
        msLevel,
        target_msLevel,
      score,
        target_name,
        matched_peaks_count
    )
```


save the resulting object:

```{r}
save(qtof_matchesF, file = here("data/database_QTOF_neg", "qtof_matchesF.RData"))
```
