---
title: "workflow to test the similarity_RDynlib() function on ftms data"
format: html
  server: shiny
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This quarto document represents the workflow for calling the `similarity_RDynlib()` from the [similarity_functions.R](https://github.com/rebdau/RDynLib_SQL/blob/Ahlam/R/similarity_functions.R) file to calculate the similarity between FTMS flax data and the RDynLib FTMS database.

# Required packages

This project follows an R package layout as follows:

The main project folder is:

RDynLib_SQL/

Within this directory, the structure is organized as follows:

RDynLib_SQL/ ├── data/ ├── R/ └── vignettes/

### `data/`

Contains the SQLite database files used in the workflow (e.g., `ftms_flax_neg.sqlite`).

### `R/`

Contains the R source files (`.R`) that define the package functions and core implementation.

### `vignettes/`

Contains Quarto (`.qmd`) files that document the workflow and provide reproducible examples.

In this workflow, the working directory is set to the project root:

RDynLib_SQL/

## Loading required packages

First, install the new CompoundDb version:

```{r}
BiocManager::install("RforMassSpectrometry/CompoundDb", ref = "jomain", force = TRUE)
```

```{r}
library(here)
library(Spectra)
library(CompoundDb)
library(MetaboAnnotation)
library(DBI)
library(RSQLite)
library(devtools)
library(dplyr)
library(stringr)
#library(RDynLib)
#` Uploading RDynLibSQL: Load all functions of the RDynlib directory
load_all()
```

## convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

Next we convert the "Dynlib_FTMS_neg.sqlite" database to a spectra object

```{r}

dbf <- here("data", "DYL_flax_herba_FTMSneg.sqlite")
db <- dbConnect(SQLite(), dbf)
dy_sps <- sql_to_spectra(dbf)
```

```{r}
spectraData(dy_sps, columns = c("retention_time"))
```

Here we load the FTMS of flax seed database that contains both spectrum types : the assembled and not assembled one (it's already a spectra object),\
you can load it from: [ftmsneg_flax_sql.qmd](https://github.com/rebdau/RDynLib_ftms/blob/ahlam/ftmsneg_flax_sql.qmd)

```{r}

load(here("data","database_FTMS_neg","ftms_all.RData"))
ftms_all <- ftms_all
spectraData(ftms_all)

```
filter out all peaks that have intensity below 1% (or 0.5%) of the base peak

```{r}
ftms_flax <- filter_basepeak(ftms_all, threshold = 0.01)
```

```{r}
cbind(
  before = lengths(peaksData(ftms_all)),
  after  = lengths(peaksData(ftms_flax))
)
```


Next we calculate the similarity score between the dynlib and flax seed FTMS negative integuments.

#test the new function(where the user should enter the spectrum type and mslevel )

We start by calculating the spectral similarity between MS2 spectra:
```{r}
ftms_matches2_T <- similarity_RDynLib_match(
  ftms_flax,
   dy_sps,
  polarity_query = 0,
   polarity_target = 0,
   requirePrecursor = TRUE,
   tolerance = 0,
   ppm = 5,
   threshold = 0.9,
   fragments_resolution = "unit.resolution"
 )
```

print the similarity matrix and the number of rows (matches)
```{r}

print(ftms_matches2_T)
nrow(ftms_matches2_T) 
```

Adding names of the new matches:
```{r}
add_spectral_matches(db, ftms_matches2_T)
```

```{r}
dbGetQuery(db, "Select * from msms_spectrum where match_name IS not null LIMIT 1")
```

look for a given name
```{r}
ftms_matches2_T %>%
  filter(str_detect(target_name, regex("herbacetin 3,8-di-o-beta-d-glucopyranoside ID AM", ignore_case = TRUE)))
```

```{r}
quercetin <- ftms_matches2_T %>%
  filter(str_detect(target_name, regex("diglucoside", ignore_case = TRUE)))

quercetin$target_name
```

Select the first 10 matches with the highest common peaks
```{r}
best_hits2 <- ftms_matches2_T %>%
    arrange(desc(matched_peaks_count)) %>%   
    slice_head(n = 10)                

best_hits
```

```{r}
best_hits2 %>%
    select(
        
        precursorMz,
        target_precursorMz,
        msLevel,
        target_msLevel,
      score,
        target_name,
        matched_peaks_count
    )
```

Next, we calculate the spectral similarity between MS3 spectra:
```{r}
ftms_matches3_T <- similarity_RDynLib_match(
  ftms_flax,
   dy_sps,
  polarity_query = 0,
   polarity_target = 0,
   requirePrecursor = TRUE,
   tolerance = 0,
   ppm = 5,
   threshold = 0.9,
   fragments_resolution = "unit.resolution"
 )
```

print the similarity matrix and the number of rows (matches)
```{r}

print(ftms_matches3_T)
nrow(ftms_matches3_T) 
```

look for a given name
```{r}
ftms_matches3_T %>%
  filter(str_detect(target_name, regex("herbacetin 3,8-di-o-beta-d-glucopyranoside ID AM", ignore_case = TRUE)))
```

Adding names of the new matches:
```{r}
add_spectral_matches(db, ftms_matches3_T)
```

```{r}
dbGetQuery(db, "Select * from msms_spectrum where match_name IS not null")
```

the 10 highest peak for ms3 matches
```{r}
best_hits3_T <- ftms_matches3_T %>%
    arrange(desc(matched_peaks_count)) %>%   
    slice_head(n = 10)                

best_hits
```


```{r}
best_hits3_T %>%
    select(
        
        precursorMz,
        target_precursorMz,
        msLevel,
        target_msLevel,
      score,
        target_name,
        matched_peaks_count
    )
```

save the resulting object:

```{r}
#DYNLIB MS3 (INCLUDING HERBA) VS FLAX
save(ftms_matches3_T, file = here("data/database_FTMS_neg", "ftms_matches3_T.RData"))
#DYNLIB MS2 (INCLUDING HERBA) VS FLAX
save(ftms_matches2_T, file = here("data/database_FTMS_neg", "ftms_matches2_T.RData"))
```

# Export the spectra that have a match

```{r}
msms_matches <- dbGetQuery(
  db,
  "
  SELECT *
  FROM msms_spectrum
  WHERE match_name IS NOT NULL
  "
)
nrow(msms_matches)
```

Export it to a csv file
```{r}
write.csv(
  msms_matches,
  "data/msms_spectrum_matches_ftneg.csv",
  row.names = FALSE
)
```


