---
title: "Untitled"
format: html
server: shiny
---

## Shiny Documents

This Quarto document is made interactive using Shiny. Interactive documents allow readers to modify parameters and see the results immediately. Learn more about Shiny interactive documents at <https://quarto.org/docs/interactive/shiny/>.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change. This demonstrates how a standard R plot can be made interactive:

```{r}
#| message: false

#' Load required libraries
library(devtools)
library(DBI)
library(RSQLite)
#rm(list = c("convert_compid_assoc"))
#` Uploading RDynLib: Load all functions of the RDynlib directory
load_all()
```

setting the arguments for MSnspecplot_assoc_SQL() function:

```{r}
#directory to the sql databases
sql_dir <- "C:/Users/ament/Desktop/sql_data"

#path to the file with the aligned compounds
assoc_path <- "C:/Users/ament/Desktop/sql_data/assoc_flax.txt"

#compound_id
dbkey <- 101171

#database name
db_name <- "DYL_FLAX_FTMSneg.sqlite"
```

Plot the spectra of a given compound and the corrosponding compounds 
in the "assoc_path".

```{r}
test <- MSnspecplot_assoc_SQL(sql_dir, db_name, dbkey, assoc_path, nr_col = 35, 
                      nr_col2 = 6, lc.err = 0.02, mz.err = 0.001, MS1 = NULL,
                      prcx = 0.6) 
```
The output of 'MSnspecplot_assoc_SQL()' shows that the 45816 is the aligned 
compound from Qtofneg with 101171 compound from FTMSneg.

Plot just the spectra of the 45816 compound from Qtof database:

```{r}
MSnplot_SQL("C:/Users/ament/Desktop/sql_data/DYL_FLAX_QTOFneg.sqlite", 45816)
```

Plot just the spectra of the 101171 compound from FTMS database:
```{r}
MSnplot_SQL("C:/Users/ament/Desktop/sql_data/DYL_FLAX_FTMSneg.sqlite", 101171)
```

Test the MS2plot to plot just the MS2 spectrum
```{r} 
MS2plot_SQL("C:/Users/ament/Desktop/sql_data/DYL_FLAX_FTMSneg.sqlite", 101171)

```


test cspp.display function
```{r}
cspp.display_SQL("C:/Users/ament/Desktop/sql_data/DYL_FLAX_FTMSneg.sqlite", 1, nr_col = 35) 
```

test gnps.display function
```{r}
gnps.display_SQL("C:/Users/ament/Desktop/sql_data/DYL_FLAX_QTOFneg.sqlite", 26, nr_col2 = 6) 

```

test MSMSplot_SQL function

```{r}
prdion <- read.table(
  "C:/Users/ament/Desktop/sql_data/product_ion_candid.txt",
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

neutloss <- read.table(
  "C:/Users/ament/Desktop/sql_data/neutral_loss_candid.txt",
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)


MSMSplot_SQL("C:/Users/ament/Desktop/sql_data/DYL_FLAX_QTOFneg.sqlite", 26,
             prdion = prdion, neutloss = neutloss, err = NULL, minum = 2, 
             oldpar = NULL, nl = NULL)
```


```{r}
MSMSspecplot_SQL("C:/Users/ament/Desktop/sql_data/DYL_FLAX_QTOFneg.sqlite", 26, 
                 prdion = prdion, neutloss = neutloss, err = NULL, 
                 minum = NULL, nr_col = NULL,   nr_col2 = NULL)
```


```{r}
#for flax ftms the conversion works
cspp_ft <- "C:/Users/ament/Desktop/sql_data/cspp_ftneg.txt"
ft_path <- "C:/Users/ament/Desktop/sql_data/DYL_FLAX_FTMSneg.sqlite"
cspp.tot_SQL(sql_path = ft_path, 266, mzerr = 0.015,
                         cspp = cspp_ft, peakwidth = NULL, IntThres  = 5)
```

```{r}
ft_con <- dbConnect(SQLite(), ft_path)
dbGetQuery(ft_con, "
  SELECT *
  FROM compound_add
  ORDER BY rowid DESC
  LIMIT 10
")

```


```{r}
#for qtof flax the conversions does not match
cspp_qtof <- "C:/Users/ament/Desktop/sql_data/cspp_qtofneg.txt"
qtof_path <- "C:/Users/ament/Desktop/sql_data/DYL_FLAX_QTOFneg.sqlite"
cspp.tot_SQL(sql_path = qtof_path, 52, mzerr = 0.015,
                         cspp = cspp_qtof, peakwidth = NULL, IntThres  = 5)

```

```{r}
#verification
DBI::dbGetQuery(con_qtof, "
  SELECT g.*
  FROM compound_add g
  INNER JOIN ms_compound c 
      ON g.compound_id = c.compound_id
  WHERE c.expid = 52
    AND (
         g.GNPS IS NOT NULL OR
         g.CSPP_RED IS NOT NULL OR
         g.CSPP_ROX IS NOT NULL OR
         g.CSPP_MET IS NOT NULL OR
         g.CSPP_OXY IS NOT NULL OR
         g.CSPP_HYD IS NOT NULL OR
         g.CSPP_BOX IS NOT NULL OR
         g.CSPP_MOX IS NOT NULL OR
         g.CSPP_ACE IS NOT NULL OR
         g.CSPP_CAR IS NOT NULL OR
         g.CSPP_PRE IS NOT NULL OR
         g.CSPP_GLY IS NOT NULL OR
         g.CSPP_MON IS NOT NULL OR
         g.CSPP_HdN IS NOT NULL OR
         g.CSPP_HBA IS NOT NULL OR
         g.CSPP_PEN IS NOT NULL OR
         g.CSPP_COU IS NOT NULL OR
         g.CSPP_RHA IS NOT NULL OR
         g.CSPP_HuN_C IS NOT NULL OR
         g.CSPP_VAN IS NOT NULL OR
         g.CSPP_HaN_C IS NOT NULL OR
         g.CSPP_HEX IS NOT NULL OR
         g.CSPP_DCA IS NOT NULL OR
         g.CSPP_HUN IS NOT NULL OR
         g.CSPP_URO IS NOT NULL OR
         g.CSPP_FER IS NOT NULL OR
         g.CSPP_GuN_C IS NOT NULL OR
         g.CSPP_SYR IS NOT NULL OR
         g.CSPP_GaN_C IS NOT NULL OR
         g.CSPP_GUN IS NOT NULL OR
         g.CSPP_SIN IS NOT NULL OR
         g.CSPP_SuN_C IS NOT NULL OR
         g.CSPP_SaN_C IS NOT NULL OR
         g.CSPP_SUN IS NOT NULL
        )
  ORDER BY g.compound_id
")
```


```{r}

MultExp_CSPP_SQL("C:/Users/ament/Desktop/sql_data/DYL_FLAX_QTOFneg.sqlite", 
                 startexp = 1, stopexp = 5, peakwidth = NULL, mzerr = 0.015,
                             IntThres  = 5, cspp = cspp)
```

Test gnps conversion
```{r}
#qtof flax
qtof_path <- "C:/Users/ament/Desktop/sql_data/DYL_FLAX_QTOFneg.sqlite"
conv.GNPS_SQL(sql_path = qtof_path, expid= 52, mzerr = 0.015,
              min = 5, adduct = 46.0055, thr1 = 3, thr2 = 0.1, thr3 = 0.4,
              IntThres = 5)
                          
```

```{r}
con_qtof <- DBI::dbConnect(RSQLite::SQLite(), qtof_path)
#show the latest compounds (the gnps rows of the flax compounds are all NA)
DBI::dbGetQuery(con_qtof, "
  SELECT *
  FROM gnps_add
  ORDER BY rowid DESC
  LIMIT 10
")

#verify the latest compound experiment
DBI::dbGetQuery(con_qtof, "
   SELECT *
  FROM ms_compound
   WHERE compound_id = 45887
 ")
DBI::dbGetQuery(con_qtof, "
  SELECT g.*
  FROM gnps_add g
  INNER JOIN ms_compound c 
      ON g.compound_id = c.compound_id
  WHERE c.expid = 52
    AND (
         g.GNPS_1 IS NOT NULL OR
         g.GNPS_2 IS NOT NULL OR
         g.GNPS_3 IS NOT NULL OR
         g.GNPS_4 IS NOT NULL OR
         g.GNPS_5 IS NOT NULL
        )
  ORDER BY g.compound_id
")

```

```{r}
#ftms flax
ft_path <- "C:/Users/ament/Desktop/sql_data/DYL_FLAX_FTMSneg.sqlite"
conv.GNPS_SQL(sql_path = ft_path, expid= 266, mzerr = 0.015,
              min = 5, adduct = 46.0055, thr1 = 3, thr2 = 0.1, thr3 = 0.4,
              IntThres = 5)
                          
```

```{r}
DBI::dbGetQuery(ft_con, "
  SELECT g.*
  FROM gnps_add g
  INNER JOIN ms_compound c 
      ON g.compound_id = c.compound_id
  WHERE c.expid = 266
    AND (
         g.GNPS_1 IS NOT NULL OR
         g.GNPS_2 IS NOT NULL OR
         g.GNPS_3 IS NOT NULL OR
         g.GNPS_4 IS NOT NULL OR
         g.GNPS_5 IS NOT NULL
        )
  ORDER BY g.compound_id
")

#verify the latest compound experiment
DBI::dbGetQuery(ft_con, "
   SELECT *
  FROM ms_compound
   WHERE compound_id = 100944 
 ")

```

GNPS and CSPP Network

```{r}
#Qtof dynlib
Overall.net_SQL(sql_path = qtof_path, exp.id = 1, dbkey = 26, min = 5,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```

```{r}
#ftms Dynlib
Overall.net_SQL(sql_path = ft_path, exp.id = 1, dbkey = 8, min = 100,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```

```{r}
#Qtof flax
qtof_path <- "C:/Users/ament/Desktop/sql_data/DYL_FLAX_QTOFneg.sqlite"
Overall.net_SQL(sql_path = qtof_path, exp.id = 52, dbkey = 45821 , min = 5,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```

```{r}
#ftms flax   
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 100915, min = 100,
                            nr_of_seq = 2,
                            thr1 = 1,
                            thr2 = 0.1,
                            thr3 = 0.2)
```



```{r}
net.lst1 <- net.addit_SQL(sql_path = sql_path, exp.id= 1, nettype="CSPP", thr1=0, thr2=0, thr3=0)
edges.net <- net.lst1[[2]]
head(edges.net)
```

```{r}
net.lst1 <- net.addit_SQL(sql_path = sql_path, exp.id = 1, nettype = "CSPP", thr1 = 0, thr2 = 0, thr3 = 0)
net.lst1
```

test MultMatchCID_SQL() function

```{r}
MultMatchCID_SQL(qtof_path, 45887, minIons = 1)
```

