---
title: "Untitled"
format: html
server: shiny
---

## Shiny Documents

This Quarto document is made interactive using Shiny. Interactive documents allow readers to modify parameters and see the results immediately. Learn more about Shiny interactive documents at <https://quarto.org/docs/interactive/shiny/>.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change. This demonstrates how a standard R plot can be made interactive:

```{r}
library(DBI)
library(RSQLite)
library(dplyr)
library(devtools)
library(here)
#library(RDynlib)
load_all()
```

#FTMS

```{r}
FT_path <- here("data", "DYL_flax_herba_FTMSneg.sqlite")

FT_con <- DBI::dbConnect(RSQLite::SQLite(), FT_path)


```

```{r}
DBI::dbExecute(
  FT_con,
  "UPDATE msms_spectrum
   SET spectrum_type = 'individual'
   WHERE spectrum_type IS 'not_assembled'"
)

```

```{r}
DBI::dbExecute(
  FT_con,
  "UPDATE msms_spectrum
   SET spectrum_type = 'pseudo_MS2'
   WHERE spectrum_type IS 'assembled'"
)
```
```{r}
dbGetQuery(FT_con, "Select * from msms_spectrum where spectrum_type = 'pseudo_MS2' limit 5" )
```
```{r}
dbGetQuery(FT_con, "Select * from msms_spectrum where spectrum_type = 'individual' limit 5" )
```
#QTOF

```{r}
QTOF_path <- "C:/Users/ament/Desktop/Github/RDynLib_SQL/data/DYL_FLAX_QTOFneg.sqlite"
QTOF_con <- DBI::dbConnect(RSQLite::SQLite(), QTOF_path)
  
```

```{r}
DBI::dbExecute(
  QTOF_con,
  "UPDATE msms_spectrum
   SET spectrum_type = 'individual'
   WHERE spectrum_type IS 'not_assembled'"
)

```
```{r}
DBI::dbExecute(
  QTOF_con,
  "UPDATE msms_spectrum
   SET spectrum_type = 'same_energy'
   WHERE spectrum_type IS 'intersect_single_energy'"
)
```

```{r}
DBI::dbExecute(
  QTOF_con,
  "UPDATE msms_spectrum
   SET spectrum_type = 'all_energies'
   WHERE spectrum_type IS 'merged_MSMS_all_energies'"
)
```

```{r}
available_spectrum_types(QTOF_con)
```

```{r}
available_spectrum_types(FT_con)
```

