---
title: "Integrated Workflow for Molecular Network Characterization, Spectral
        Annotation, and Database Curation"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This Quarto document presents a workflow for the characterization and 
visualization of molecular networks. It includes the analysis of known
compounds detected in FT-MS data and their corresponding features in QTOF 
datasets, as well as the visualization of ion trap–based and QTOF-based neutral
loss analyses. The document further summarizes the results of CID spectral 
matching, describes the annotation and incorporation of characterized molecules 
into the database, and outlines the conversion of the global SQL database into
an MGF format compatible with public repositories and analysis platforms 
for further comparison with external public spectral databases.

## Load required libraries

```{r}
#| message: false
#| results: "hide"
#| warning: false
#' Load required libraries
library(devtools)
library(DBI)
library(RSQLite)
library(here)
library(CompoundDb)
library(MsBackendMgf)
#rm(list = c("convert_compid_assoc"))
#` Uploading RDynLib: Load all functions of the RDynlib directory
load_all()
```
#FTMS flax negative
```{r}
#| message: false
#| results: "hide"
#| warning: false
ft_path <- here("data", "DYL_FLAX_FTMSneg.sqlite")
con <- dbConnect(RSQLite::SQLite(), ft_path)
cspp_ft <- here("data","cspp_flax.txt")
charge_file <- here("data", "ftms_all_compounds_RD.xlsx")
```


We start by the CID spectral matching of 101280 compound from ftms negative 
database, To do that we need to convert tha database to a spectra object:
```{r}
dy_sps <- sql_to_spectra(ft_path)
```


```{r}
cmp <- MultMatchCID_Spectra(
  dy_sps,
  101280,
   spectrum_type_query = "not_assembled",
  spectrum_type_target = "NA",
   ms_level_query = 2,
   ms_level_target = 2,
  polarity_query = 0,
  polarity_target = 0,
  tolerance = 0,
  ppm = 10,
  minIons = ,
    fragments_resolution = "unit.resolution",
  requirePrecursor = T,
)
head(cmp, 5)
```

Then we update the cspp table:
```{r}
#| message: false
#| results: "hide"
#| warning: false
cspp.tot_SQL(ft_path, 266, mzerr = 0.01,
                         cspp = cspp_ft, peakwidth = 0.1, IntThres  = 0, mz_tol = 0.5,
                         dot_thresh = 0, spectrum_type = "assembled",  
                         min_common = 0, charge_file = charge_file)
```

Update the gnps table:
```{r}
#| message: false
#| results: "hide"
#| warning: false
conv.GNPS_SQL(sql_path= ft_path, expid = 266, peakwidth = 0.2, mzerr = 0.015,
              min = 5, adduct = 60.0211, thr1 = 2, thr2 = 0.1, thr3 = 0.3,
              spectrum_type = "not_assembled", IntThres = 0)
```

Next we visualize the CSPP and GNPS network of 101280 compound from ftms neg:
```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 20
#| fig-height: 15
Overall.net_SQL(sql_path = ft_path, exp.id = 266, dbkey = 101280, min = 0,
                nr_of_seq = 3, thr1 = 0.1, thr2 = 0.2, thr3 = 0.3)
```

Now we show the ion trap based analysis of 100988 compound:

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSnplot_SQL(ft_path, 100988, thr_int = 100)
```

```{r}
#| fig-width: 14
#| fig-height: 4
MSnplot_SQL(ft_path, 101155, thr_int = 100)
```

```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSnplot_SQL(ft_path, 101283, thr_int = 100)
```


#Qtof flax analysis

Import the database and required files:
```{r}
qt_path <- here("data", "DYL_FLAX_QTOFneg.sqlite")
con_qt <- dbConnect(RSQLite::SQLite(), qt_path )
neutral_loss_candid <- here("data", "neutral_loss_candid.txt")
neutloss <- read.table(
  neutral_loss_candid,
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

product_ion_candid <- here("data", "product_ion_candid_2.txt")
prdion <- read.table(
  product_ion_candid,
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

```


The aligned compound in qtof data that corresponds to 100988 
compound from ftms data:
```{r}
alignment_file <- here("data", "assoc_flax_63.txt")

getAlignedCompids <- getAlignedCompids_SQL(
compound_id = 100988,
  database_name = "DYL_FLAX_FTMSneg.sqlite",
  alignment_file = alignment_file
)
getAlignedCompids
```
The 45890 is the qtof compound_id that was aligned with 100988 ftms compound_id. 

The msms_spectrum table linked to that compound_id to select the spectrum_id:
```{r}
list_spectrum <- get_spectrum_id(45890)
list_spectrum
```

We chose the 46001 spectrum_id to show the qtof based neutral loss analysis:
```{r}
#| message: false
#| results: "hide"
#| warning: false
#| #| dev: svg
#| fig-width: 14
#| fig-height: 8
MSMSspecplot_SQL(qt_path, 46001, 
                 prdion = prdion, neutloss = neutloss, err = NULL, 
                 minum = NULL, nr_col = NULL,   nr_col2 = NULL)
```

Now we enter the characterization of the compounds found to the database:
```{r}
#EnterName_SQL(ft_path, CoMPID = 101143)
```


```{r}
#EnterName_SQL(ft_path, CoMPID = 100988)
```


Next, we call the function to export the sql database to an mgf format that 
could be used for further analysis and comparisons with public databases 
available on public platforms like: GNPS...
We start by converting ftms data
```{r}
#| message: false
#| results: "hide"
#| warning: false
ft_assembled <- here("data", "ft_assembled.mgf")
sql_to_mgf(ft_path, expid = 266, spectrum_type = "assembled", 
           output_file = ft_assembled) 
```

Then, we convert qtof data:
```{r}
#| message: false
#| results: "hide"
#| warning: false
qt_assembled <- here("data", "qt_assembled.mgf")
sql_to_mgf(qt_path, expid = 52, spectrum_type = "intersect_single_energy", 
           output_file = qt_assembled) 
```




