---
title: "SQL, Compdb, and Spectra workflow"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

# Shiny Documents

This Quarto document demonstrates
(1) how to subset a DynLib SQL database by including certain experiments
(2) how the DynLib infrastructure interoperates with the R for Mass Spectrometry ecosystem.
(3) how to use RDynLib Accessors for Compound and Spectral Data

# Load required libraries

```{r}
#| message: false
#| results: "hide"
#| warning: false
library(here)
library(Spectra)
library(CompoundDb)
library(DBI)
library(RSQLite)
library(RDynLib)
library(devtools)
load_all()
```
# Subset a DynLib SQL database
The DynLib SQL database can archive many LC-MS experiments. The expid of each experiment can be retrieved in the experiment table. Here we filter an SQL database to contain only a single experiment experiment.

COMMENT: we need to put an example dynlib database in the package, consisting of a subset of a published maize experiment and our new flax experiment

## File paths to the input and output sql databases
COMMENT: why output ?
```{r}
sql_db <- here("data", "DYL_FLAX_FTMSneg.sqlite")
```
## retrieve the expid
COMMENT: put here SQL command to extract experiment table from the database

We identify the expid of the experiment of interest: expid = ....

## subsetting
```{r}
filterDynLibDB(sql_db, expid=)
```

# Convert a DynLib database to CompDb and Spectra objects
Both `CompDb` and `Spectra` objects are part of the provide access to the R for Mass Spectrometry ecosystem. While `CompDb` offers a database-centric interface for querying and managing compound database content,`Spectra` provides a more spectrum-centric interface.

## CompDb object
We create a `CompDb` object `cdb` from the DynLib SQL database using the `ExpCOmpDb()` function from the `RDynLib` package. The resulting `CompDb` object represents the complete SQL database and provides access to all of its tables.
```{r}
cdb <- ExpCompDb(sql_db)
# Print database metadada information
meta <- metadata(cdb)
head(meta)
str(cdb)
```


## Spectra object
A `CompDb` object can be converted into a `Spectra` object using the `Spectra()` function from the `Spectra` package. By default, the MzBackendCompDb backend is used, which keeps the spectra linked to the underlying compound database, allowing the entire database to be accessed through the resulting `Spectra` object.
Alternatively, `RDynLib` provides a function to convert a DynLib SQL database directly into a Spectra object with MzBackendCompDb backend. 

```{r}
# Using the Spectra package:
sps <- Spectra(cdb)
sps

# Using the RDynLib package:
sps.rdl <- sql_to_spectra(sql_db)

str(sps)
str(sps.rdl)


```
The objects `sps` and `sps.rdl` are exactly the same `Spectra` objects.

Print spectra variables:
```{r}
spectraVariables(sps)
```

Print spectra Data:
```{r}
spectraData(sps, columns = c("msLevel" , "retention_time", "precursorMz",
                                "centroided","scanIndex", "dataOrigin", 
                                "compound_id", "nodename" 
                                ))
```

All the columns from the original SQL table are preserved and reported in 
the spectraData of the `Spectra` object.
# Using RDynLib Accessors for Compound and Spectral Data
#To facilitate retrieval of compound information in such cases,  \texttt{RDynLib} provides dedicated shortcut functions that exploit information available for all detected compounds. The  #\texttt{NodenameMass\_SQL()} function enables retrieval of compound records from a DynLib database based on a nodename, a complete or partial compound name, or a measured precursor m/z range within one or more #selected experiments, thereby providing the corresponding  \texttt{compound\_id} required for subsequent access to spectral data. The  \texttt{get\_spectrum\_id()} function retrieves the associated  #\texttt{spectrum\_id}(s) and spectral metadata from the  \texttt{msms\_spectrum} table, enabling extraction of the corresponding MS/MS or MS$^n$ spectra.

# Retrieving a compound_id in a DynLib SQL database, using the `CompoundDb` package.
# maybe not necessary
Compound information can be extracted from a `CompDb` object using the `compounds()` function from the `CompoundDb` package. The resulting table corresponds to the `ms_compound` table in the SQL database and contains the compound identifier (compound_id), together with chromatographic information such as m/z and retention time, and structural annotation data, such as the compound name and molecular formula. By applying filters in the `compounds()` function, one can create a subsetted ms_compund table with compound identifiers (compound_id) of selected compounds corresponding to a defined value for `name` or `exactmass`.

```{r}
cmp <- compounds(cdb)
head(cmp)
```

The compound information of specific criteria:
```{r}
supportedFilters(cdb)
compounds(cdb, columns = c("compound_id", "name", "formula", "expid"),
          filter = ~ name == "quercetin-hexoside-deoxyhexoside KA KM")

```

# Retrieving a spectrum of a specific compound from a DynLib SQL database, using the `CompoundDb` and `Spectra` packages.

Spectral data can be accessed through a `Spectra` object. A spectrum can be retrieved by first filtering the `Spectra` object for a compound of interest using its `compound_id`. The corresponding spectrum identifiers (`spectrum_id`) (one or more may exist per compound_id) can then be extracted and used to select a specific spectrum. Finally, the fragment *m/z* and intensity values can be retrieved from the selected spectrum.

Here we retrieve a spectrum for `compound_id ==2`.

```{r}
## Filter the Spectra object by compound_id
## Retrieve spectrum_id corresponding to a compound_id.
## Filter Spectra object based on spectrum_id.
## Retrieve spectrum m/z intensity list.

```




COMMENT: This tables() command doesn't work for me:
```{r}
tables(cdb)
```




For more queries and details visit this workflow: https://www.bioconductor.org/packages//release/bioc/vignettes/CompoundDb/inst/doc/CompoundDb-usage.html




