---
title: "Integrating SQL in RDynlib workflow"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

# Shiny Documents

This Quarto document demonstrates
(1) how to subset a DynLib SQL database by including certain experiments.
(2) how the DynLib infrastructure interoperates with the R for Mass Spectrometry
    ecosystem.
(3) how to use RDynLib Accessors for Compound and Spectral Data.

# Load required libraries

```{r}
#| message: false
#| results: "hide"
#| warning: false
library(here)
library(Spectra)
library(CompoundDb)
library(DBI)
library(RSQLite)
#library(RDynLib)
library(devtools)
load_all()
```
# Subset a DynLib SQL database
The DynLib SQL database can archive many LC-MS experiments. The expid of each experiment can be retrieved in the experiment table. Here we filter an SQL database to contain only a single experiment experiment.

## File paths to the input and output sql databases
COMMENT: why output ?
Answer : is for storing the resulting sql database that contains just the selected experiment. 
```{r}
sql_db <- here("data", "DYL_FLAX_FTMSneg.sqlite")
output_db <- here("data", "subset.sqlite")
```

## retrieve the expid

```{r}
con <- dbConnect(RSQLite::SQLite(), sql_db)

#Verify the experiment table
dbGetQuery(con, "
SELECT *
FROM experiment LIMIT 5
")
```

## Call the function to filter the database based on a single expid
```{r}
filterDynLibDB(
  expid = 266,
  input_db = sql_db,
  output_db = output_db
)
```

## Connection to the subseted database
```{r}
con_filter <- dbConnect(RSQLite::SQLite(), output_db)

#Verify the experiment table
dbGetQuery(con_filter , "
SELECT *
FROM experiment LIMIT 5
")
```

## Print ms_compound table of the filtered database
```{r}
dbGetQuery(con_filter , "
SELECT *
FROM ms_compound LIMIT 5
")
```

## filter the database based on a set of expids
```{r}
output_db2 <- here("data", "filter_set.sqlite")
filterDynLibDB(
    expid = c(5, 7, 12),
    input_db = sql_db,
    output_db = output_db2
)
```

## Connection to the subseted database
```{r}
con_filter2 <- dbConnect(RSQLite::SQLite(), output_db2)

#Verify the experiment table
dbGetQuery(con_filter2 , "
SELECT *
FROM experiment
")
```

# Convert a DynLib database to CompDb and Spectra objects
Both `CompDb` and `Spectra` objects are part of the provide access to the R for Mass Spectrometry ecosystem. While `CompDb` offers a database-centric interface for querying and managing compound database content,`Spectra` provides a more spectrum-centric interface.

## CompDb object
We create a `CompDb` object `cdb` from the DynLib SQL database using the `ExpCompDb()` function from the `RDynLib` package. The resulting `CompDb` object represents the complete SQL database and provides access to all of its tables.
```{r}
cdb <- ExpCompDb(sql_db)
# Print database metadada information
meta <- metadata(cdb)
head(meta)
str(cdb)
```

## Retrieving a compound_id in a DynLib SQL database, using the `CompoundDb` package.
Compound information can be extracted from a `CompDb` object using the `compounds()` function from the `CompoundDb` package. The resulting table corresponds to the `ms_compound` table in the SQL database and contains the compound identifier (compound_id), together with chromatographic information such as m/z and retention time, and structural annotation data, such as the compound name and molecular formula. By applying filters in the `compounds()` function, one can create a subsetted ms_compund table with compound identifiers (compound_id) of selected compounds corresponding to a defined value for `name` or `exactmass`.

```{r}
cmp <- compounds(cdb)
head(cmp)
```

The compound information of specific criteria:
```{r}
supportedFilters(cdb)
compounds(cdb, columns = c("compound_id", "name", "formula", "expid"),
          filter = ~ name == "quercetin-hexoside-deoxyhexoside KA KM")

```

For more queries and details visit this workflow: https://www.bioconductor.org/packages//release/bioc/vignettes/CompoundDb/inst/doc/CompoundDb-usage.html

## Spectra object
A `CompDb` object can be converted into a `Spectra` object using the `Spectra()` function from the `Spectra` package. By default, the MzBackendCompDb backend is used, which keeps the spectra linked to the underlying compound database, allowing the entire database to be accessed through the resulting `Spectra` object.
Alternatively, `RDynLib` provides a function to convert a DynLib SQL database directly into a Spectra object with MzBackendCompDb backend. 

```{r}
# Using the Spectra package:
sps <- Spectra(cdb)
sps

# Using the RDynLib package:
sps.rdl <- sql_to_spectra(sql_db)

str(sps)
str(sps.rdl)


```
The objects `sps` and `sps.rdl` are exactly the same `Spectra` objects.

Print spectra variables:
```{r}
spectraVariables(sps)
```

Print spectra Data:
```{r}
spectraData(sps, columns = c("msLevel" , "retention_time", "precursorMz",
                                "centroided","scanIndex", "dataOrigin", 
                                "compound_id", "nodename" 
                                ))
```

All the columns from the original SQL table are preserved and reported in 
the spectraData of the `Spectra` object.
# RDynLib Accessors for Compound and Spectral Data
To facilitate retrieval of compound information in such cases,  \texttt{RDynLib} provides dedicated shortcut functions that exploit information available for all detected compounds. The  #\texttt{NodenameMass\_SQL()} function enables retrieval of compound records from a DynLib database based on a nodename, a complete or partial compound name, or a measured precursor m/z range within one or more #selected experiments, thereby providing the corresponding  \texttt{compound\_id} required for subsequent access to spectral data. The  \texttt{get\_spectrum\_id()} function retrieves the associated  \texttt{spectrum\_id}(s) and spectral metadata from the  \texttt{msms\_spectrum} table, enabling extraction of the corresponding MS/MS or MS$^n$ spectra.

testing the NodenameMass_SQL() function that extract the additional cololumns 
and rows from ms_compound table of a given experiment, precursor_mz or name 
or nodename.

Here we test using expid and precursor mz within a mass range:
```{r}
FT_path <- here("data", "DYL_FLAX_FTMSneg.sqlite")

NodenameMass_SQL(
  sql_path = sql_db,
  expid = 1,
  mass_measured = 537.1976,
  mass_range = 0.02
)
```

Here we test using expid and a part of the name of the compounds we want to show:
```{r}
res <-NodenameMass_SQL(
  sql_path = sql_db,
  expid = 1,
  name = "hexoside"
)
head(res,3)
```

show the msms_spectrum table of a given compound_id
```{r}
sp_id <- get_spectrum_id(qt_path = sql_db, compound_id = 101280 )
sp_id
```

Retreive the ms_compound table of a given compound_id
```{r}
compound <- get_compound(qt_path = sql_db, compound_id = 101280)
compound
```
