---
title: "SQL, Compdb, and spectra workflow"
format:
  html:
    toc: true
    self-contained: true
author: "Ahlam Mentag, Rebecca Dauwe, Johannes Rainer"
---

## Shiny Documents

This Quarto document is made to demonstrate how to convert an sql database to 
a CompDB object and then to a spectra object.

## Load required libraries

```{r}
#| message: false
#| results: "hide"
#| warning: false
library(here)
library(Spectra)
library(CompoundDb)
library(DBI)
library(RSQLite)
library(RDynLib)
load_all()
```
#File paths to the input and output sql databases
```{r}
sql_db <- here("data", "DYL_FLAX_FTMSneg.sqlite")
```

#Convert SQl to a CompDb object
```{r}
cdb <- ExpCompDb(sql_db)
```

Print Compounds information
```{r}
cmp <- compounds(cdb)
head(cmp)
```

```{r}
tables(cdb)
```

The compound information of specific criterea:
```{r}
compounds(cdb, columns = c("compound_id", "name", "formula"),
          filter = ~ name == "quercetin-hexoside-deoxyhexoside KA KM")
```

Print metadada information
```{r}
meta <- metadata(cdb)
head(meta)
```
For more queries and details visit this workflow: https://www.bioconductor.org/packages//release/bioc/vignettes/CompoundDb/inst/doc/CompoundDb-usage.html


#Convert the compDB object to a spectra object:
```{r}
sps <- Spectra(cdb)
sps
```
Print spectra variables:
```{r}
spectraVariables(dy_sps)
```

Print specta Data:
```{r}
spectraData(dy_sps, columns = c("msLevel" , "retention_time", "precursorMz",
                                "centroided","scanIndex", "dataOrigin", 
                                "compound_id", "nodename" 
                                ))
```

All the columns from the original SQL table are preserved and reported in 
the spectraData of the converted specta object.


